    
\chapter{同音の漢字による書き換え}

\begin{center}
\begin{Large}
第363課: 同音の漢字による書き換え 
\end{Large}
\end{center}
 
\par{ The practice of 同音の漢字による書き換え (rewriting with homophonous Kanji) began in the year 1956 by the Japanese Language Council (国語審議会) when it announced a directive for rewriting Kanji with homophonous counterparts when said Kanji were not a part of the 当用漢字表 established by the same committee. This practice also refers to the standardization of spellings for which there was more than one Kanji spelling available. }

\par{ In 1981, however, the 当用漢字表 was replaced with the 常用漢字表, and with the replacement, 書き換え lost some of its compelling force. This was not solely due to switching to the newly revised character list for general use. In fact, the proliferation of word processors alone was causing the need for such a practice to diminish rapidly.  Its effect, however, remains as a component of style guides for official use. Some words have forever had their primary spellings changed and may never revert back to what they are. In this lesson, you will be able to see the most common examples of 書き換え. With knowledge of prior spellings, many words will make more sense and you'll have less difficulty reading things written before 1956. }
      
\section{The Ironic Complication of Simplification}
 
\par{ A lot of factors have contributed to the mess in orthography 書き換え has caused. One such factor was the concept of 代用字, characters chosen to unilaterally replace another character. This is the main contributing factor to examples aside from unifying spellings. }

\par{ An extremely important thing to note as you read through the many examples presented in this lesson is how much of an impact the practice of 書き換え has had on both the written and spoken language. You will notice that the overwhelming majority of examples are exemplary of the written language. A lot of words that had their spellings altered simply fell out of use. In fact, only about 40\% of the words listed in the charts can be classified as common even in the spoken language. }

\par{\textbf{Orthography Note }: Spellings used in example sentences are stilted toward using replaced characters. This is because the traditional spelling(s) of many of these words have fallen almost entirely out of use. However, if you wish to substitute newer spellings with older ones, you may do so but at the risk of the reader potentially not being able to read the word. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Recitation & あんしょう & 暗誦・諳誦 & 暗唱 & Dark night & あんや & 闇夜 & 暗夜 \\ \cline{1-8}

Intention & いこう & 意嚮 & 意向 & Reparation & いしゃりょう & 慰藉料 & 慰謝料 \\ \cline{1-8}

Garment & いしょう & 衣裳 & 衣装 & Gloom & いんえい & 陰翳 & 陰影 \\ \cline{1-8}

Genius & えいさい & 穎才 & 英才 & Wisdom & えいち & 叡智 & 英知 \\ \cline{1-8}

Protection & えんご & 掩護 & 援護 & Favor & おんぎ & 恩誼 & 恩義 \\ \cline{1-8}

\end{ltabulary}

\par{1. 今学期より ${\overset{\textnormal{かんしあんしょう}}{\text{漢詩暗誦}}}$ スキルという授業が開かれます。 \hfill\break
A new course called "Chinese Poem Recitation"will be given starting this semester. }

\par{2. 去年のハロウィンで、ナチスの ${\overset{\textnormal{ぐんぷく}}{\text{軍服}}}$ を ${\overset{\textnormal{れんそう}}{\text{連想}}}$ させるとされる ${\overset{\textnormal{いしょう}}{\text{衣装}}}$ を着たアイドルグループがイスラエルから ${\overset{\textnormal{こうぎ}}{\text{抗議}}}$ を受けるというニュースがあった。 \hfill\break
There was a news report in last year's Halloween about a pop idol group which wore outfits reminiscent of Nazi military uniforms. }

\par{3. ${\overset{\textnormal{げんしばくだんひばくしゃ}}{\text{原子爆弾被爆者}}}$ を ${\overset{\textnormal{えんご}}{\text{援護}}}$ すること。 \hfill\break
Helping atomic bomb victims. }

\par{4. 厳しい ${\overset{\textnormal{にゅうがくしけん}}{\text{入学試験}}}$ に合格して入学してくるほど英才であれば、そうした学生はすでに大学教育に ${\overset{\textnormal{ふさわ}}{\text{相応}}}$ しい学力を持っているでしょう。 \hfill\break
If someone is the kind of genius who is admitted into a university after having passed a harsh entry exam, that student already has the academic ability becoming of a college education. }

\par{5. ${\overset{\textnormal{ひゆ}}{\text{比喩}}}$ や ${\overset{\textnormal{いんえい}}{\text{陰影}}}$ に ${\overset{\textnormal{と}}{\text{富}}}$ んだ文章だからこそ、読者の方々の心に響かせるでしょう。 \hfill\break
Because it is a such a prose filled with metaphors and subtleties, it is sure to reverberate in the minds of readers. }

\par{6. 古代の ${\overset{\textnormal{えいち}}{\text{叡智}}}$ は素晴らしいなと思っています。 \hfill\break
I think that ancient wisdom is amazing. }

\par{7. 長い暗夜にも必ず夜明けがある。 \hfill\break
Dawn will always break even for the longest of dark nights. }

\par{8. ${\overset{\textnormal{すで}}{\text{既}}}$ に ${\overset{\textnormal{ししゅつ}}{\text{支出}}}$ した ${\overset{\textnormal{ちりょうひ}}{\text{治療費}}}$ と今後かかるであろう治療費を含めてこれから ${\overset{\textnormal{かがいしゃ}}{\text{加害者}}}$ に ${\overset{\textnormal{いしゃりょう}}{\text{慰謝料}}}$ を求めようと思っています。 \hfill\break
I intend to hereafter seek reparation from the perpetrator for medical expenditures I've already incurred as well as those I expect to incur in the future. }

\par{\textbf{Controversy Note }: 援護 and 掩護 are technically different words. 援護 refers to rescuing those in trouble whereas 掩護 refers to protecting allies from enemy attack. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Rotation & かいてん & 廻転 & 回転 & Blaze & かえん & 火焰 & 火炎 \\ \cline{1-8}

Scuffling & かくとう & 挌闘 & 格闘 & Lively & かっぱつ & 活潑 & 活発 \\ \cline{1-8}

Intermittent & かんけつ & 間歇 & 間欠 & Vital & かんじん & 肝腎 & 肝心 \\ \cline{1-8}

Scarce & きしょう & 稀少 & 希少 & Badge & きしょう & 徽章 & 記章 \\ \cline{1-8}

Miracle & きせき & 奇蹟 & 奇跡 & Colorful story & きだん & 綺談 & 奇談 \\ \cline{1-8}

Counseling & きょうかい & 教誨 & 教戒 & Lethal weapon & きょうき & 兇器 & 凶器 \\ \cline{1-8}

\end{ltabulary}
 
\par{\textbf{Controversy Note }: 記章 and 徽章 are technically different words. 記章 refers to an insignia used to commemorate something. 徽章 refers to badge indicating position\slash qualifications\slash affiliation. }

\par{\textbf{Controversy Note }: 奇談 and 綺談 are technically different words. 奇談 refers to strange stories. 綺談 refers to well-crafted stories. }

\par{\textbf{Controversy Note }: 教戒 and 教誨 are technically different words. 教戒 refers to admonishing. 教誨 refers to doing pastoral work to prisoners. }

\par{9. 格闘技を習いたいと思っているのですが、なかなかどれを選べばいいのか分かりません。 \hfill\break
I'd like to study martial arts, but I'm not quite sure which one I should choose. }
 
\par{10. 肝心なところをしっかりと理解してほしい。 \hfill\break
I want you to firmly understand the essentials. }
 
\par{11. 家が腫れ上げる火炎に飲まれ瞬く間に爆発した。 \hfill\break
The home was swallowed in bellowing flames and exploded in an instant. }
 
\par{12. 彼は見ての通り元気で活発な少年です。 \hfill\break
As you can see, he is a lively and active youth. }
 
\par{13. 度重なる奇蹟に驚きました。 \hfill\break
I was surprised at the repeated miracles. }
 
\par{14. もし凶器を持って「殺すぞ」と脅した時、相手が「やってみな」などと言い返したとします。 \hfill\break
Let's suppose that you threaten someone by saying "I'll kill you" while holding a deadly weapon and your opponent retorts back "I'd like to see you try." }

\par{15. 貴子は、ここまで長々と話を聞かされた以上、偶然でも奇跡でも構わないから、その美容師の記憶力に間違いがないことを祈るだけだと思いながら、ひたすら黙って立っていた。 \hfill\break
Seeing that she had been made to listen to this long-winded talk, Takako earnestly stood there silently as she felt there was no other recourse but to pray that there was no doubt in the stylist's memory, whether it be by sheer coincidence or by miracle. \hfill\break
From 凍える牙 by 乃南アサ. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Fishing & ぎょろう & 漁撈 & 漁労 \hfill\break
& Talent & ぎりょう & 伎倆 \hfill\break
& 技量 \\ \cline{1-8}

Compartment & くかく & 区劃 & 区画 & Excavation & くっさく & 掘鑿 & 掘削 \\ \cline{1-8}

Admonition & くんかい & 訓誡 & 訓戒 & Smoking (food) & くんせい & 燻製 & 薫製 \\ \cline{1-8}

Washout & けっかい & 決潰 & 決壊 & Rising to action & けっき & 蹶起 & 決起 \\ \cline{1-8}

Farewell & けつべつ & 訣別 & 決別 & String music & げんか & 絃歌 & 弦歌 \\ \cline{1-8}

\end{ltabulary}

\par{16. ${\overset{\textnormal{ぎょろうせつび}}{\text{漁労設備}}}$ が船ごとに異なるのは仕方がない。 \hfill\break
It can't be helped that fishing equipment differs per bus. }

\par{17. ${\overset{\textnormal{ぎりょう}}{\text{技量}}}$ を ${\overset{\textnormal{みが}}{\text{磨}}}$ くために ${\overset{\textnormal{さら}}{\text{更}}}$ なる ${\overset{\textnormal{けんしゅう}}{\text{研修}}}$ が必要だ。 \hfill\break
Further training is needed to improve skills. }

\par{18. 海外の ${\overset{\textnormal{ぎょうせいくかく}}{\text{行政区画}}}$ を「 ${\overset{\textnormal{けん}}{\text{県}}}$ 」と ${\overset{\textnormal{やく}}{\text{訳}}}$ す ${\overset{\textnormal{きじゅん}}{\text{基準}}}$ は何でしょうか。 \hfill\break
What is the standard for translating overseas administrative districts as "prefectures"? }

\par{19. ${\overset{\textnormal{きぬがわ}}{\text{鬼怒川}}}$ の ${\overset{\textnormal{ぼうはてい}}{\text{防波堤}}}$ が ${\overset{\textnormal{けっかい}}{\text{決壊}}}$ し、家が次々と ${\overset{\textnormal{だくりゅう}}{\text{濁流}}}$ に飲まれてしまった。 \hfill\break
The Kinugawa seawall broke and house after house was swallowed up by a muddy stream. }

\par{20. ${\overset{\textnormal{こきょう}}{\text{故郷}}}$ に ${\overset{\textnormal{けつべつ}}{\text{決別}}}$ し、自分なりの思いの地で ${\overset{\textnormal{せいかつさいけん}}{\text{生活再建}}}$ に ${\overset{\textnormal{せんしん}}{\text{専心}}}$ したいと思う。 \hfill\break
Saying farewell to my hometown, I'd like to devote myself to resettling akin to my own mind. }

\par{21. 3本の ${\overset{\textnormal{うんが}}{\text{運河}}}$ の ${\overset{\textnormal{くっさくりょう}}{\text{掘削量}}}$ を減らすために、 ${\overset{\textnormal{いく}}{\text{幾}}}$ つかの ${\overset{\textnormal{じんぞうこ}}{\text{人造湖}}}$ を作る計画が出された。 \hfill\break
A plan to construct several man-made lakes was brought up to reduce the amount of excavation work for the three canals. }

\par{22. ${\overset{\textnormal{とうふ}}{\text{豆腐}}}$ の ${\overset{\textnormal{くんせい}}{\text{燻製}}}$ に ${\overset{\textnormal{ちょうせん}}{\text{挑戦}}}$ しました。 \hfill\break
I challenged myself with smoking tofu. }

\par{23. 政府に ${\overset{\textnormal{さくしゅ}}{\text{搾取}}}$ され続けてきた国民が ${\overset{\textnormal{けっき}}{\text{決起}}}$ して ${\overset{\textnormal{よとうかんぶ}}{\text{与党幹部}}}$ らを ${\overset{\textnormal{だとう}}{\text{打倒}}}$ することに成功した。 \hfill\break
The citizens, who had continued to be exploited by the government, succeeded in overthrowing the party leaders by rising to action. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Solemn & げんぜん & 儼然 & 厳然 & Precipice & けんそ & 嶮岨 & 険阻 \\ \cline{1-8}

Crossing & こうさ & 交叉 & 交差 & Deduction & こうじょ & 扣除 & 控除 \\ \cline{1-8}

Ore & こうせき & 礦石 & 鉱石 & Funeral offering & こうでん & 香奠 & 香典 \\ \cline{1-8}

Far-reaching & こうはん & 広汎 & 広範 & Excitement & こうふん & 亢奮・昂奮 & 興奮 \\ \cline{1-8}

Public relations & こうほう & 弘報 & 広報 & Reconciliation & こうわ & 媾和 & 講和 \\ \cline{1-8}

\end{ltabulary}

\par{24. ${\overset{\textnormal{ばくやく}}{\text{爆薬}}}$ を ${\overset{\textnormal{せっち}}{\text{設置}}}$ して ${\overset{\textnormal{こうざん}}{\text{鉱山}}}$ を ${\overset{\textnormal{ばくは}}{\text{爆破}}}$ することで ${\overset{\textnormal{こうせき}}{\text{鉱石}}}$ を ${\overset{\textnormal{ほ}}{\text{掘}}}$ り ${\overset{\textnormal{お}}{\text{起}}}$ こすことができる。 \hfill\break
You can dig up ore by setting off dynamite in a mine. }

\par{25. 日本は 1951 年 9 月 8 日 にサンフランシス ${\overset{\textnormal{こうわじょうやく}}{\text{講和条約}}}$ を ${\overset{\textnormal{ていけつ}}{\text{締結}}}$ し ${\overset{\textnormal{せんりょう}}{\text{占領}}}$ から ${\overset{\textnormal{かいほう}}{\text{開放}}}$ され ${\overset{\textnormal{せいしき}}{\text{正式}}}$ に ${\overset{\textnormal{しゅうせん}}{\text{終戦}}}$ となりました。 \hfill\break
Japan entered into the Treaty of San Francisco on September 8th, 1951, which released Japan from occupation and officially ended the war. }

\par{26. この道の ${\overset{\textnormal{とくしょく}}{\text{特色}}}$ のひとつは ${\overset{\textnormal{ひかくてき}}{\text{比較的}}}$ に ${\overset{\textnormal{けんそ}}{\text{険阻}}}$ な ${\overset{\textnormal{さかみち}}{\text{坂道}}}$ が多いことである。 \hfill\break
One characteristic of this road is that there is a lot of relatively steep slopes. }

\par{27. ${\overset{\textnormal{しょとくぜいこうじょ}}{\text{所得税控除}}}$ を受けるためには、 ${\overset{\textnormal{かくていしんこく}}{\text{確定申告}}}$ の手続きを行う必要があります。 \hfill\break
In order to receive an income tax credit, having a final tax form done is necessary. }

\par{28. ${\overset{\textnormal{けいえい}}{\text{経営}}}$ に ${\overset{\textnormal{かん}}{\text{関}}}$ する ${\overset{\textnormal{こうはん}}{\text{広範}}}$ な知識を身につける使命を持っているんです。 \hfill\break
I have a mission to acquire far-reaching knowledge in management. }

\par{29. ヒトの ${\overset{\textnormal{こっかく}}{\text{骨格}}}$ とチンパンジーの骨格と比べてみましょう。  \hfill\break
Let's compare the skeletal structures of humans and chimpanzees. }

\par{30. 自分と ${\overset{\textnormal{たいさ}}{\text{大差}}}$ ないと思う、きっかけ一つ間違えただけだと思うのに、貴子と彼女たちとの間には、 ${\overset{\textnormal{げんぜん}}{\text{厳然}}}$ たる違いがあった。 \hfill\break
Although Takako felt they weren't so different from herself, and although she felt that they had only messed up on one chance, there was a solemn difference between them and her. \hfill\break
From 凍える牙 by 乃南アサ . }

\par{31. 駅に向かって歩きながら、貴子の中で奇妙な ${\overset{\textnormal{こうふん}}{\text{興奮}}}$ が ${\overset{\textnormal{わ}}{\text{湧}}}$ き起こっていた。 \hfill\break
As Takako walked toward the train station, a curious excitement welled up inside her. \hfill\break
From 凍える牙 by 乃南アサ . }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Running dry & こかつ & 涸渇 & 枯渇 & Physique\slash framework & こっかく & 骨骼 & 骨格 \\ \cline{1-8}

Employment & こよう & 雇傭 & 雇用 & Root\slash basis & こんてい & 根柢 & 根底 \\ \cline{1-8}

Acetic acid & さくさん & 醋酸 & 酢酸 & Turmoil & こんめい & 昏迷 & 混迷 \\ \cline{1-8}

Idle spectating & ざし & 坐視 & 座視 & Traffic jam & ざっとう & 雑沓 & 雑踏 \\ \cline{1-8}

Compliment\slash eulogy & さんじ & 讃辞 & 賛辞 & Irrigating & さんすい & 撒水 & 散水 \\ \cline{1-8}

Stimulus & しげき & 刺戟 & 刺激 & Corpse & したい & 屍体 & 死体 \\ \cline{1-8}

\end{ltabulary}

\par{\textbf{Controversy Note }: 混迷 and 昏迷 are technically different words. 混迷 refers to being bewildered with no insight to be had. 昏迷 refers to being stupefied, or in medical terms, being in a coma. }

\par{32. 世界の ${\overset{\textnormal{しげん}}{\text{資源}}}$ が ${\overset{\textnormal{こかつ}}{\text{枯渇}}}$ しつつある。 \hfill\break
The world's resources are running dry. }

\par{33. ${\overset{\textnormal{こようしょうめいしょ}}{\text{雇用証明書}}}$ を ${\overset{\textnormal{ていしゅつ}}{\text{提出}}}$ してください。 \hfill\break
 Please submit employment verification. }

\par{34. ${\overset{\textnormal{ざっとうたいさく}}{\text{雑踏対策}}}$ が ${\overset{\textnormal{けいし}}{\text{軽視}}}$ されていた。 \hfill\break
Congestion countermeasures had been neglected. }

\par{35. ${\overset{\textnormal{こうじげんば}}{\text{工事現場}}}$ で ${\overset{\textnormal{さんすいしゃ}}{\text{散水車}}}$ を運転する。 \hfill\break
 To drive a sprinkler trunk at a construction site. }

\par{36. 死んだ後も、朽ちることのない死体と ${\overset{\textnormal{か}}{\text{化}}}$ してしまったのだった。  \hfill\break
Even after death, he transformed into a corpse which experienced no decay. }

\par{37. あの ${\overset{\textnormal{ころ}}{\text{頃}}}$ は、刑事ごっこをしているようで、新たな刺激に包まれつつも、何とも気楽なものだった。  \hfill\break
At that time, it seemed like (she) was playing detective,and while she was surrounded with new stimuli, she was quite at ease. \hfill\break
From 凍える牙 by 乃南アサ . }

\par{38. 自分が今まで信じ、 ${\overset{\textnormal{ささ}}{\text{支}}}$ えにしていたものは ${\overset{\textnormal{こんてい}}{\text{根底}}}$ から ${\overset{\textnormal{くつがえ}}{\text{覆}}}$ されてしまった。 \hfill\break
What I had believed and supported was overturned from the ground up. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Wheeled vehicle & しゃりょう & 車輛 & 車両 & Collection & しゅうしゅう & 蒐集 & 収集 \\ \cline{1-8}

Being resolved & しゅうそく & 終熄 & 終息 & Settlement & しゅうらく & 聚落 & 集落 \\ \cline{1-8}

Impediment & しょうがい & 障碍・障礙 & 障害 & Effacement & しょうきゃく & 銷却 & 消却 \\ \cline{1-8}

Advancement & しょうじょ & 陞叙 & 昇叙 & Impatience & しょうそう & 焦躁 & 焦燥 \\ \cline{1-8}

Barrier & しょうへき & 牆壁 & 障壁 & Distillation & じょうりゅう & 蒸溜 & 蒸留 \\ \cline{1-8}

\end{ltabulary}
 
\par{\textbf{Controversy Note }: 収集 and 蒐集 are technically different words. 収集 refers to people\slash things "gathering" and 蒐集 refers to collecting lots of things\slash material out of interest\slash for study. }

\par{\textbf{Controversy Note }: 障壁 and 牆壁 are technically different words. 障壁 refers to an obstruction. 牆壁 refers to fences and walls. }

\par{39. ${\overset{\textnormal{こうがくしゃりょう}}{\text{高額車輛}}}$ を ${\overset{\textnormal{ゆそう}}{\text{輸送}}}$ する場合に輸送コストは ${\overset{\textnormal{わりだか}}{\text{割高}}}$ になってしまうが、セキュリティー上の理由から利用されることが多い。 \hfill\break
Transporting high-cost rail cars means transport costs are comparatively high, but they are used a lot for security reasons. }

\par{40. ${\overset{\textnormal{いしき}}{\text{意識}}}$ して ${\overset{\textnormal{まえむ}}{\text{前向}}}$ きに ${\overset{\textnormal{しこう}}{\text{思考}}}$ することが必要なのは不安と ${\overset{\textnormal{きょうふ}}{\text{恐怖}}}$ と ${\overset{\textnormal{しょうそうかん}}{\text{焦燥感}}}$ に ${\overset{\textnormal{か}}{\text{駆}}}$ られているときなどでしょう。 \hfill\break
It is necessary to consciously think positively whenever one is frantic with anxiety, fear, and impatience. }

\par{41. 韓国の ${\overset{\textnormal{ぱく}}{\text{朴}}}$ 大統領は ${\overset{\textnormal{いあんふもんだい}}{\text{慰安婦問題}}}$ が日本との ${\overset{\textnormal{がいこう}}{\text{外交}}}$ の ${\overset{\textnormal{しょうへき}}{\text{障壁}}}$ になっていると発言した。 \hfill\break
South Korea President Park remarked that the issue of comfort women was an obstacle in diplomacy with Japan. }

\par{42. ${\overset{\textnormal{めいぼ}}{\text{名簿}}}$ から名前を ${\overset{\textnormal{しょうきゃく}}{\text{消却}}}$ する。 \hfill\break
 To erase one's name from the register. }

\par{43. ${\overset{\textnormal{しゅうらく}}{\text{集落}}}$ が ${\overset{\textnormal{ぜんしょう}}{\text{全焼}}}$ し 200 人が犠牲となった。  \hfill\break
The whole settlement burned down, resulting in 200 victims. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Epistle & しょかん & 書翰 & 書簡 & Lyricism & じょじょう & 抒情 & 叙情 \\ \cline{1-8}

Trial & しれん & 試煉 & 試練 & Corrosion & しんしょく & 浸蝕 & 浸食 \\ \cline{1-8}

Stretching & しんちょう & 伸暢 & 伸長 & Permeation & しんとう & 滲透 & 浸透 \\ \cline{1-8}

Interrogation & じんもん & 訊問 & 尋問 & Degeneration & すいたい & 衰頽 & 衰退 \\ \cline{1-8}

Inhabiting & せいそく & 棲息 & 生息 & (Leading) edge & せんたん & 尖端 & 先端 \\ \cline{1-8}

Acute\slash radical & せんえい & 尖鋭 & 先鋭 & Screening & せんこう & 銓衡 & 選考 \\ \cline{1-8}

Cleaning & せんじょう & 洗滌 & 洗浄 & Sensationalism & せんじょう & 煽情 & 扇情 \\ \cline{1-8}

\end{ltabulary}

\par{\textbf{Controversy Note }: 洗浄 and 洗滌 are technically different words. 洗浄 refers to cleansing in a religious sense. 洗滌 refers to cleaning with water\slash medicine. }

\par{\textbf{Controversy Note }: 生息 and 棲息 are technically different words. 生息 refers either to "living one's life" or plants living. 棲息 refers either to (animal) habitation. }

\par{\textbf{Controversy Note }: 先端 and 尖端 are technically different words. 先端 refers to the end of something long. 尖端 refers either to the tip of something sharp or the start of a trend\slash period. }

\par{44. ${\overset{\textnormal{ゆうびんしょかん}}{\text{郵便書簡}}}$ とは ${\overset{\textnormal{ふうとうけんよう}}{\text{封筒兼用}}}$ の ${\overset{\textnormal{びんせん}}{\text{便箋}}}$ のことです。 \hfill\break
A letter-card is letter paper combined with an envelope. }

\par{45. 最初の ${\overset{\textnormal{しれん}}{\text{試練}}}$ にチャレンジしたが失敗に終わった。 \hfill\break
I challenged the first trial, but I ended in failure. }

\par{46. 中国は21世紀になっても膨張主義で ${\overset{\textnormal{りんこく}}{\text{隣国}}}$ を次々と ${\overset{\textnormal{しんしょく}}{\text{侵食}}}$ していった。 \hfill\break
Even into the 21 st century, China eroded into its neighboring countries one after another under expansionism. }

\par{47. 沖縄に対する ${\overset{\textnormal{へんこう}}{\text{偏向}}}$ が日本社会に ${\overset{\textnormal{しんとう}}{\text{浸透}}}$ している。 \hfill\break
A bias toward Okinawa has percolated into Japanese society. }

\par{48. ${\overset{\textnormal{せんじょうてき}}{\text{扇情的}}}$ な見出しの記事のほうが読者を増やすのが当たり前だ。 \hfill\break
It\textquotesingle s only natural that sensational headlines are what increases readers. }

\par{49. ${\overset{\textnormal{せんじょうざい}}{\text{洗浄剤}}}$ や ${\overset{\textnormal{えんそけい}}{\text{塩素系}}}$ ${\overset{\textnormal{ひょうはくざい}}{\text{漂白剤}}}$ などを誤って飲んだ時、すぐに水で口をすすいでから、コップ1～2杯の牛乳または水を飲んでください。 \hfill\break
When you accidentally ingest detergent or chlorine bleach, immediately rinse your mouth with water and then drink one or two cups of milk or water. }

\par{\textbf{漢字Note }: すすぐ in the meaning of “to rinse out (one\textquotesingle s mouth)” can be spelled as 漱ぐ. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Arbitrariness & せんだん & 擅断 & 専断 & Death in battle & せんぼつ & 戦歿 & 戦没 \\ \cline{1-8}

Inlaying & ぞうがん & 象嵌 & 象眼 & Fellowship & じょうぎ & 情誼 & 情義 \\ \cline{1-8}

Integration & そうごう & 綜合 & 総合 & Rivalry & そうこく & 相剋 & 相克 \\ \cline{1-8}

Annihilation & そうめつ & 剿滅 & 掃滅 & Growing in clusters & そうせい & 簇生 & 族生 \\ \cline{1-8}

Hindrance & そし & 沮止 & 阻止 & Communication & そつう & 疏通 & 疎通 \\ \cline{1-8}

Faded color & たいしょく & 褪色 \hfill\break
& 退色 & Entreaty & たんがん & 歎願 & 嘆願 \\ \cline{1-8}

Indoor heating & だんぼう & 煖房 & 暖房 & Wisdom\slash wit & ちえ & 智慧 & 知恵 \\ \cline{1-8}

Knowledge & ちしき & 智識 & 知識 & Remark\slash annotation & ちゅうしゃく & 註釈 & 注釈 \\ \cline{1-8}

\end{ltabulary}

\par{\textbf{Controversy Note }: 情義 and 情誼 are technically different words. 情義 refers to humanity in a sense of duty. 情誼 refers to cordiality with others. }
 
\par{\textbf{Reading Note }: 族生・簇生 may alternatively be read as ぞくせい. }

\par{50. 様々な生物の ${\overset{\textnormal{せいそくぶんぷ}}{\text{棲息分布}}}$ が大きく変わってしまったら生態系が段々と壊れていくのだろう。 \hfill\break
If the habitats of various creatures were to greatly change, the ecosystem would likely gradually collapse. }
 
\par{51. ${\overset{\textnormal{ろじょう}}{\text{路上}}}$ で私服警官に ${\overset{\textnormal{じんもん}}{\text{尋問}}}$ を受ける。 \hfill\break
To be questioned by a plain‐clothes officer on the street. }
 
\par{52. この部分の ${\overset{\textnormal{しんちょう}}{\text{伸長}}}$ は、 ${\overset{\textnormal{どじょうちゅう}}{\text{土壌中}}}$ でのみ起こり、光が当たると ${\overset{\textnormal{ていし}}{\text{停止}}}$ してしまう。 \hfill\break
The elongated growth of this area only occurs when it is in soil and halts whenever light hits it. }
 
\par{53. ${\overset{\textnormal{せいふしゅさい}}{\text{政府主催}}}$ の ${\overset{\textnormal{ぜんこくせんぼつしゃついとうしき}}{\text{全国戦没者追悼式}}}$ が行われた。 \hfill\break
A government sponsored national memorial service for war dead was held. }
 
\par{54. 職務上の ${\overset{\textnormal{けんげん}}{\text{権限}}}$ を ${\overset{\textnormal{こ}}{\text{越}}}$ えた ${\overset{\textnormal{せんだんてき}}{\text{専断的}}}$ な行為をしてはならない。 \hfill\break
One mustn't commit arbitrary acts that go beyond the authority of one's work duties. }
 
\par{55. ${\overset{\textnormal{せんさい}}{\text{繊細}}}$ な ${\overset{\textnormal{きんぞくぞうがん}}{\text{金属象嵌}}}$ で ${\overset{\textnormal{くさばな}}{\text{草花}}}$ の ${\overset{\textnormal{もよう}}{\text{模様}}}$ を表す。 \hfill\break
To represent a flower design with delicate metal inlay. }
 
\par{56. 義理と人情の ${\overset{\textnormal{そうこく}}{\text{相克}}}$ に ${\overset{\textnormal{なや}}{\text{悩}}}$ む。 \hfill\break
To be distressed over the rivalry between duty and humanity. }
 
\par{57. 周りと ${\overset{\textnormal{いしそつう}}{\text{意思疎通}}}$ ができなくても、身振り手振りで何となく分かり合えるようになるのではないかと思います。 \hfill\break
I think that even if you couldn't mutually understand with those around you that you would somehow be able to understand each other with gestures. }
 
\par{58. 暖房を付けると電気代がぐんと上がってしまう。 \hfill\break
One's electric bill drastically goes up when the heater is turned on. }
 
\par{59. 心理学の知識を生かす仕事って何でしょう。 \hfill\break
What are some jobs that utilize knowledge in psychology? }
 
\par{60. Wordでの注釈の付け方がわかりません。 \hfill\break
I don't know how to add annotations in Word. }

\par{61. 決して ${\overset{\textnormal{しんぽ}}{\text{進歩}}}$ を ${\overset{\textnormal{そし}}{\text{阻止}}}$ するためではない。 \hfill\break
It is never to obstruct progress. }

\par{62. ${\overset{\textnormal{しんにゅうしゃ}}{\text{侵入者}}}$ を ${\overset{\textnormal{そうめつ}}{\text{掃滅}}}$ せよ。 \hfill\break
Annihilate intruders. }

\par{63. ${\overset{\textnormal{せんもんてき}}{\text{専門的}}}$ かつ ${\overset{\textnormal{そうごうてき}}{\text{総合的}}}$ な ${\overset{\textnormal{いけん}}{\text{意見}}}$ を ${\overset{\textnormal{と}}{\text{取}}}$ り ${\overset{\textnormal{まと}}{\text{纏}}}$ めました。 \hfill\break
I have compiled expert and comprehensive opinion (on the matter). }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Order & ちゅうもん & 註文 & 註文 & Strange occurrence & ちんじ & 椿事 & 珍事 \\ \cline{1-8}

Precipitation & ちんでん & 沈澱 & 沈殿 & Collision & ていしょく & 牴触・觝触 & 抵触 \\ \cline{1-8}

Courteous & ていちょう & 鄭重 & 丁重 & Polite & ていねい & 叮嚀 & 丁寧 \\ \cline{1-8}

Anchorage & ていはく & 碇泊 & 停泊 & Notebook & てちょう & 手帖 & 手帳 \\ \cline{1-8}

Tumbling & てんとう & 顚倒 & 転倒 & Emulation & とうしゅう & 蹈襲 & 踏襲 \\ \cline{1-8}

Stoppage & とぜつ & 杜絶 & 途絶 & Immorality & はいとく & 悖徳 & 背徳 \\ \cline{1-8}

Revocation & はき & 破毀 & 破棄 & Exposure\slash disclosure & ばくろ & 曝露 & 暴露 \\ \cline{1-8}

Crushing & はさい & 破摧 & 破砕 & Fermentation & はっこう & 醱酵 & 発酵 \\ \cline{1-8}

Excerpt & ばっすい & 抜萃 & 抜粋 & Insurrection & はんらん & 叛乱 & 反乱 \\ \cline{1-8}

\end{ltabulary}
 
\par{\textbf{Controversy Note }: 珍事 and 椿事 are technically different words. 珍事 refers to a rare event. 椿事 refers to unexpected occurrence usually with a negative nuance. }

\par{\textbf{Controversy Note }: 破棄 and 破毀 are technically different words. 破棄 refers to ridding of an agreement\slash contract. 破毀 refers to repealing a ruling upon appeal. }

\par{\textbf{Meaning Note }: 沈殿・沈澱 means "precipitation as in the formation of a solid in chemistry. }
 
\par{64. ${\overset{\textnormal{しゃくほうたんがんしょ}}{\text{釈放嘆願書}}}$ に ${\overset{\textnormal{しょめい}}{\text{署名}}}$ する。 \hfill\break
To sign an acquittal petition. }
 
\par{65. 化学の ${\overset{\textnormal{じっけんそうさほう}}{\text{実験操作法}}}$ として ${\overset{\textnormal{ちんでん}}{\text{沈殿}}}$ が利用される場合がある。 \hfill\break
There are instances in which precipitation is used as a chemical experimental procedure. }
 
\par{66. 一人一人のお客さんと丁寧に向き合って仕事をすることが大事です。 \hfill\break
It is important to work while facing each and every customer politely. }
 
\par{67. ${\overset{\textnormal{てんとう}}{\text{転倒}}}$ しない限り、ガソリンが出てくることはないはずです。 \hfill\break
So long as you don't tumble (in it), there shouldn't ever be gasoline coming out of it. }
 
\par{68. 米軍の ${\overset{\textnormal{はんざい}}{\text{犯罪}}}$ を ${\overset{\textnormal{ばくろ}}{\text{暴露}}}$ したマニング ${\overset{\textnormal{いっ}}{\text{一}}}$ ${\overset{\textnormal{とうへい}}{\text{等兵}}}$ が20の ${\overset{\textnormal{ようぎ}}{\text{容疑}}}$ について有罪の ${\overset{\textnormal{ひょうけつ}}{\text{評決}}}$ を受けた。 \hfill\break
Private First Class Officer Manning received a guilty verdict on twenty counts for exposing US military crimes. }
 
\par{69. ${\overset{\textnormal{はいとくこうい}}{\text{背徳行為}}}$ をしなくても ${\overset{\textnormal{ざいあくかん}}{\text{罪悪感}}}$ を持つことはあります。 \hfill\break
There are times I have a sense of guilt even when I don't do something immoral. }
 
\par{70. 各地で反乱が起きている。 \hfill\break
Revolts are happening everywhere. }
 
\par{71. ワインはブドウを ${\overset{\textnormal{はっこう}}{\text{発酵}}}$ させて作ります。 \hfill\break
Wine is made by fermenting grapes. }
 
\par{72. 気に入っているフレーズを ${\overset{\textnormal{ばっすい}}{\text{抜粋}}}$ して一冊の本にする。 \hfill\break
To turn phrases one likes and has extracted into a book. }
 
\par{73. お手数ですがメールを破棄してくださいますようお願いいたします。 \hfill\break
We apologize for the trouble, but we ask that you would please destroy e-mails. }
 
\par{74. ノースウェスト ${\overset{\textnormal{じゅんしゅう}}{\text{準州}}}$ に ${\overset{\textnormal{てんざい}}{\text{点在}}}$ する町から町への ${\overset{\textnormal{こうくうゆそう}}{\text{航空輸送}}}$ が ${\overset{\textnormal{とぜつ}}{\text{途絶}}}$ したら、多くの人が ${\overset{\textnormal{こりつじょう}}{\text{孤立状}}}$ 態になってしまい、 ${\overset{\textnormal{しんこく}}{\text{深刻}}}$ な ${\overset{\textnormal{しょくりょうぶそく}}{\text{食糧不足}}}$ に ${\overset{\textnormal{ちょくめん}}{\text{直面}}}$ せざるを得ないだろう。 \hfill\break
If air transport between town and town dotting the Northwest Territories were suspended, many people would be isolated and would have to face severe food shortage. }
 
\par{75. トランプ大統領は ${\overset{\textnormal{しゅうにん}}{\text{就任}}}$ したら間もなく、オバマ ${\overset{\textnormal{ぜんせいけん}}{\text{前政権}}}$ の ${\overset{\textnormal{ほうしん}}{\text{方針}}}$ を ${\overset{\textnormal{とうしゅう}}{\text{踏襲}}}$ しない ${\overset{\textnormal{いこう}}{\text{意向}}}$ を ${\overset{\textnormal{せいしき}}{\text{正式}}}$ に ${\overset{\textnormal{ひょうめい}}{\text{表明}}}$ した。 \hfill\break
No sooner after President Trump had taken office, he officially expressed his intent not to follow suit with Former President Obama's policies. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Baseless rumor & ひご & 蜚語 & 飛語 & Cipher\slash mark & ふちょう & 符牒 & 符丁 \\ \cline{1-8}

Compilation & へん & 篇 & 編 & Editing & へんしゅう & 編輯 & 編集 \\ \cline{1-8}

Nurturing & ほいく & 哺育 & 保育 & Obstruction & ぼうがい & 妨碍・妨礙 & 妨害 \\ \cline{1-8}

Abandonment & ほうき & 抛棄 & 放棄 \hfill\break
& Defense & ぼうぎょ & 防禦 & 防御 \\ \cline{1-8}

Bandage & ほうたい & 繃帯 & 包帯 & Huge\slash swelling & ぼうだい & 厖大 & 膨大 \\ \cline{1-8}

Kitchen knife & ほうちょう & 庖丁 & 包丁 & Adviser & ほさ & 輔佐 & 補佐 \\ \cline{1-8}

Groping for & もさく & 摸索 & 模索 & Vulgar\slash crude & やひ & 野鄙 & 野卑 \\ \cline{1-8}

Welding & ようせつ & 熔接 & 溶接 & Greed\slash desire & よく & 慾 & 欲 \\ \cline{1-8}

Theory\slash reason & りくつ & 理窟 & 理屈 & Clever & りこう & 悧巧 & 利口 \\ \cline{1-8}

Pillage & りゃくだつ & 掠奪 & 略奪 & Comprehension & りょうかい & 諒解 & 了解 \\ \cline{1-8}

Contour & りんかく & 輪廓 & 輪郭 & Linkage & れんけい & 連繫 & 連係 \\ \cline{1-8}

Alliance & れんごう & 聯合 & 連合 & Curve\slash bend & わんきょく & 彎曲 & 湾曲 \\ \cline{1-8}

\end{ltabulary}

\par{\textbf{Controversy Note }: 膨大 and 厖大 are technically different words. 膨大 refers to getting larger and is traditionally just a verb. 厖大 refers to extent\slash amount being large and is traditionally just an adjective. }

\par{\textbf{Controversy Note }: 保育 and 哺育 are technically different words. 保育 refers to raising a child\slash infant. 哺育 refers to mammals (including humans) nursing. }

\par{\textbf{Controversy Note }: 妨害 and 妨碍・妨礙 are technically different words. 妨害 refers to obstruction in a roadway. 妨碍・妨礙 refers to causing harm to someone and obstructing the execution of something. }

\par{76. インターネット上の ${\overset{\textnormal{りゅうげんひご}}{\text{流言飛語}}}$ に ${\overset{\textnormal{たい}}{\text{対}}}$ する ${\overset{\textnormal{きせい}}{\text{規制}}}$ の ${\overset{\textnormal{ぜひ}}{\text{是非}}}$ を ${\overset{\textnormal{りろんてき}}{\text{理論的}}}$ に ${\overset{\textnormal{けんとう}}{\text{検討}}}$ すべく、 ${\overset{\textnormal{とうろんかい}}{\text{討論会}}}$ が行われた。 \hfill\break
A debate took place to logically consider the pros and cons of regulations on wild groundless Internet rumors }

\par{77. ${\overset{\textnormal{つねひごろ}}{\text{常日頃}}}$ から ${\overset{\textnormal{すしや}}{\text{寿司屋}}}$ とかで『あがり』などの ${\overset{\textnormal{ふちょう}}{\text{符丁}}}$ を使う人を恥ずかしく思わないんですか。 \hfill\break
Aren\textquotesingle t you ashamed of people who always use the coded speech of venues such as sushi restaurants, for example, “agari (for tea).” }

\par{78. トランプ氏は本当にロシアの ${\overset{\textnormal{せんきょぼうがい}}{\text{選挙妨害}}}$ を認めるのか。 \hfill\break
Will Mr. Trump really admit to Russian obstruction in the election? }

\par{79. 一緒に課題を考え、一緒に解決の道を ${\overset{\textnormal{もさく}}{\text{模索}}}$ しましょう。 \hfill\break
Together, let\textquotesingle s think of tasks and fumble for ways of solving them. }

\par{80. ${\overset{\textnormal{かた}}{\text{硬}}}$ いまま ${\overset{\textnormal{むり}}{\text{無理}}}$ やり ${\overset{\textnormal{ほうちょう}}{\text{包丁}}}$ で切っていました。 \hfill\break
I was forcibly cutting it with a kitchen knife while it was still hard. }

\par{81. 犯人は、コンビニから ${\overset{\textnormal{しょうひん}}{\text{商品}}}$ を ${\overset{\textnormal{りゃくだつ}}{\text{略奪}}}$ したりパトカーに火をつけたりしていたということです。 \hfill\break
The criminal I said to have plundered a convenience store, set a police car on fire, and other things. }

\par{82. ${\overset{\textnormal{わんきょく}}{\text{湾曲}}}$ した ${\overset{\textnormal{かいがんせん}}{\text{海岸線}}}$ が見えてきた。 \hfill\break
A curved coastline came into view. }

\par{83. アラブ ${\overset{\textnormal{れんごうきょうわこく}}{\text{連合共和国}}}$ は1958年にエジプト共和国とシリア共和国と連合して作られた国家でした。 \hfill\break
The United Arab Republic was a nation created in 1958 by the unification of the Republic of Egypt and the Syrian Republic. }

\par{84. そうしたら、何が正しいのかという ${\overset{\textnormal{りせいてき}}{\text{理性的}}}$ な ${\overset{\textnormal{かんてん}}{\text{観点}}}$ に立って、つまり ${\overset{\textnormal{りくつ}}{\text{理屈}}}$ に ${\overset{\textnormal{かな}}{\text{適}}}$ った説明ができ、相手を説得することができるでしょう。 \hfill\break
And then, when you take a logical point of view of what\textquotesingle s correct, in other words, when you can make an explanation that\textquotesingle s rational, you should be able to persuade your opponent. }

\par{85. ${\overset{\textnormal{ようせつかんきょう}}{\text{溶接環境}}}$ における ${\overset{\textnormal{かんきせつび}}{\text{換気設備}}}$ のメンテナンスでは、サービス ${\overset{\textnormal{ぎじゅつしゃ}}{\text{技術者}}}$ の安全を確保する必要があります。 \hfill\break
It is necessary to secure the safety of service technicians during maintenance on the ventilation equipment in the welding environment. }

\par{86. ${\overset{\textnormal{やひ}}{\text{野卑}}}$ な ${\overset{\textnormal{ことばづか}}{\text{言葉遣}}}$ いは絶対にいけません。 \hfill\break
Vulgar speech is never tolerable. }

\par{87. トランプ氏は13日、政権の ${\overset{\textnormal{かなめ}}{\text{要}}}$ となる ${\overset{\textnormal{じきしゅせきだいとうりょうほさかん}}{\text{次期首席大統領補佐官}}}$ に ${\overset{\textnormal{きょうわとうぜんこくいいんちょう}}{\text{共和党全国委員長}}}$ のラインス・プリーバス氏を ${\overset{\textnormal{きよう}}{\text{起用}}}$ すると発表しました。 \hfill\break
On the 13th, Mr. Trump announced that he will appoint Reince Priebus, Republican National Committee Chairman, to be his administration\textquotesingle s pivotal appointee as the President Elect\textquotesingle s White House Chief of Staff. }

\par{88. 彼女は昌代の包帯を巻かれた手をさすり、「怖かったわよね」「痛かったでしょう」などと言った。 \hfill\break
She stroked the bandaged hands of Masayo as she repeatedly said, "I bet it was so scary" and "I\textquotesingle m sure it hurt." \hfill\break
From 凍える牙 by 乃南アサ. }

\par{89. つまり、今回の事件については、ホシの ${\overset{\textnormal{りんかく}}{\text{輪郭}}}$ が非常に ${\overset{\textnormal{つか}}{\text{摑}}}$ みにくいということだった。 \hfill\break
In other words, for this case, it was going to be extremely difficult to get a hold of a sketch of the perpetrator. \hfill\break
From 凍える牙 by 乃南アサ. }

\begin{center}
 \textbf{Unofficial Examples of 書き換え }
\end{center}

\par{  The Japanese Language Council, incidentally, didn't do such a good job at covering every instance of 書き換え that would be necessary to comply with their overarching initiative. As such, the community at large implemented its own simplifications, many of which are still in common use. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Dumbbell & あれい & 啞鈴 & 亜鈴 & Atrophy & いしゅく & 萎縮 & 委縮 \\ \cline{1-8}

Getting rich quick & いっかくせんきん & 一攫千金 \hfill\break
& 一獲千金 & Spoliation & いんめつ & 湮滅 & 隠滅 \\ \cline{1-8}

Conjecture & おくだん & 臆断 & 憶断 & Cowardice & おくびょう & 臆病 & 憶病 \\ \cline{1-8}

Roundworm & かいちゅう & 蛔虫 & 回虫 & Excursion & かいゆう & 回游・洄游 & 回遊 \\ \cline{1-8}

Itemization & かじょうがき & 箇条書き & 個条書き & Resolute & かっこ & 確乎 & 確固 \\ \cline{1-8}

Form\slash posture & かっこう & 恰好 & 格好 & Physique & かっぷく & 恰幅 & 格幅 \\ \cline{1-8}

Function & かんすう & 函数 & 関数 & Dignity & かんろく & 貫禄 & 貫録 \\ \cline{1-8}

Alms & ぎえん & 義捐 & 義援 & Gleeful & きき & 嬉々 & 喜々 \\ \cline{1-8}

Defamation & きそん & 毀損 & 棄損 & Drive\slash spirit & きはく & 気魄 & 気迫 \\ \cline{1-8}

Sophistry & きべん & 詭辯 & 奇弁・危弁 & Divertimenti & きゆうきょく & 嬉遊曲 & 喜遊曲 \\ \cline{1-8}

\end{ltabulary}

\par{90. ${\overset{\textnormal{きはらけいじ}}{\text{木原刑事}}}$ は ${\overset{\textnormal{みつはたようぎしゃ}}{\text{光畑容疑者}}}$ の ${\overset{\textnormal{きはく}}{\text{気魄}}}$ に押されて後ずさりした。 \hfill\break
Detective Kihara recoiled from being intimidated by the zeal of the suspect Mitsuhata. }

\par{91. これは ${\overset{\textnormal{めいよきそんざい}}{\text{名誉毀損罪}}}$ に問われる内容でしょうか。 \hfill\break
Is this the sort of stuff one can be accused of libel for? }

\par{92. 一般の角度に対する ${\overset{\textnormal{さんかくかんすう}}{\text{三角関数}}}$ を得るためには、三角関数について成り立つ何らかの定理を ${\overset{\textnormal{ししん}}{\text{指針}}}$ として、定義の拡張を行う必要がある。 \hfill\break
In order to get a trigonometric function for a general angle, it\textquotesingle s necessary that you expand its definition by using some theorem which holds as an index. }

\par{From the Wikipedia page on 三角関数. }

\par{93. 母親が ${\overset{\textnormal{いしゅく}}{\text{萎縮}}}$ したり、謝ったりする必要はありません。 \hfill\break
There is no need for mother to cower or apologize. }

\par{94. 両手に鉄の\{亜鈴・ダンベル\}を持ち上げる。 \hfill\break
To lift an iron dumbbell with both hands. }

\par{95. 猫は、 ${\overset{\textnormal{かいちゅう}}{\text{回虫}}}$ を持ったネズミや ${\overset{\textnormal{ことり}}{\text{小鳥}}}$ といった ${\overset{\textnormal{しょうどうぶつ}}{\text{小動物}}}$ を食べた場合や、 ${\overset{\textnormal{かんせん}}{\text{感染}}}$ した母猫の乳を子猫が飲んだ場合などにも、回虫に感染することがある。 \hfill\break
Cats may be infected with roundworms by eating mice, small birds, and other small animals with them and kittens may also be infected by drinking milk from their mothers if infected. }

\par{96. ${\overset{\textnormal{はんそくけんぎしゃ}}{\text{犯則嫌疑者}}}$ が ${\overset{\textnormal{とうひ}}{\text{逃避}}}$ ${\overset{\textnormal{も}}{\text{若}}}$ しくは証拠を ${\overset{\textnormal{いんめつ}}{\text{湮滅}}}$ する恐れがあって緊急を ${\overset{\textnormal{よう}}{\text{要}}}$ する場合には、事後に ${\overset{\textnormal{れいじょうこうふ}}{\text{令状交付}}}$ を ${\overset{\textnormal{せいきゅう}}{\text{請求}}}$ することができる。 \hfill\break
When there is an urgent concern that a tax evader may flee or destroy evidence, you are able to request a warrant a posteriori. }

\par{97. 東日本大震災の ${\overset{\textnormal{ふっこうしえん}}{\text{復興支援}}}$ のために ${\overset{\textnormal{ぎえんきん}}{\text{義援金}}}$ を ${\overset{\textnormal{きふ}}{\text{寄付}}}$ しました。 \hfill\break
I donated contributions for the recovery efforts for the Great East Japan Earthquake. }
 
\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Feast\slash banquet & きょうえん & 饗宴 & 供宴 & Bridgehead & きょうとうほ & 橋頭堡 & 橋頭保 \\ \cline{1-8}

Conjugation (math) & きょうやく & 共軛 & 共役 & Contribution & きょきん & 醵金 & 拠金 \\ \cline{1-8}

Contribution & きょしゅつ & 醵出 & 拠出 & Imprisonment & きんこ & 禁錮 & 禁固 \\ \cline{1-8}

Uniformity & きんせい & 均斉 & 均整 & Fumigation & くんじょう & 燻蒸 & 薫蒸 \\ \cline{1-8}

Pious & けいけん & 敬虔 & 敬謙 & Indignant & げきこう & 激昂 & 激高 \\ \cline{1-8}

Scapula\slash shoulder blade & けんこうこつ & 肩胛骨 & 肩甲骨 & Building-to-land ratio & けんぺいりつ & 建蔽率 & 建坪率 \\ \cline{1-8}

Fascination & げんわく & 眩惑 & 幻惑 & Custody\slash abduction & こういん & 勾引 & 拘引 \\ \cline{1-8}

Cultivator\slash tiller & こううんき & 耕耘機 & 耕運機 & Exacerbated & こうしん & 昂進・亢進 & 高進 \\ \cline{1-8}

Joint management & ごうべん & 合辦 & 合弁 & Harmonious & こんぜん & 渾然 & 混然 \\ \cline{1-8}

Three-forked road & さんさろ & 三叉路 & 三差路 & Article for pawning & しちぐさ & 質種 & 質草 \\ \cline{1-8}

\end{ltabulary}

\par{98. ${\overset{\textnormal{くんじょう}}{\text{薫蒸}}}$ とは ${\overset{\textnormal{ゆうがいせいぶつ}}{\text{有害生物}}}$ を ${\overset{\textnormal{くじょ}}{\text{駆除}}}$ する方法のひとつです。 \hfill\break
Fumigation is one way of exterminating pests. }

\par{99. 私のことなんですが、 ${\overset{\textnormal{けいけん}}{\text{敬虔}}}$ な ${\overset{\textnormal{むしろんしゃ}}{\text{無視論者}}}$ と言っていいかもしれません。 \hfill\break
Although this is about me, it\textquotesingle s probably safe to call me a pious atheist. }

\par{100. ${\overset{\textnormal{げっこう}}{\text{激昂}}}$ して ${\overset{\textnormal{しんぱんいん}}{\text{審判（員）}}}$ を ${\overset{\textnormal{ねら}}{\text{狙}}}$ って投げつける。 \hfill\break
To indignantly target and pelt an umpire\slash referee. }

\par{101. ${\overset{\textnormal{てっこうかんれん}}{\text{鉄鋼関連}}}$ の ${\overset{\textnormal{ごうべんじぎょうけいやく}}{\text{合弁事業契約}}}$ を結ぶ。 \hfill\break
To sign a steel-related joint venture contract. }

\par{102. ${\overset{\textnormal{じょうりくさくせん}}{\text{上陸作戦}}}$ を成功させるには、 ${\overset{\textnormal{きょうとうほ}}{\text{橋頭堡}}}$ を ${\overset{\textnormal{きず}}{\text{築}}}$ くことが ${\overset{\textnormal{ふかけつ}}{\text{不可欠}}}$ な ${\overset{\textnormal{じょうけん}}{\text{条件}}}$ である。 \hfill\break
In order to make a landing operation succeed, establishing a bridgehead is an essential condition. }

\par{103. 一度見たら忘れられない ${\overset{\textnormal{きんせい}}{\text{均整}}}$ の取れた形と色になっています。 \hfill\break
It has a balanced form and color that you won\textquotesingle t be able to forget just by looking at it once. }

\par{104. 私が借りようとしている ${\overset{\textnormal{しゃくや}}{\text{借家}}}$ ですが、 ${\overset{\textnormal{りっち}}{\text{立地}}}$ が ${\overset{\textnormal{さんさろ}}{\text{三差路}}}$ に面した場所です。 \hfill\break
Although this is about the rental home I\textquotesingle m trying to rent, it\textquotesingle s the place where the lot faces a three-forked road. }

\par{105. ${\overset{\textnormal{けんこうこつ}}{\text{肩甲骨}}}$ の筋肉を動かして刺激することで ${\overset{\textnormal{きそたいしゃ}}{\text{基礎代謝}}}$ があがる。 \hfill\break
Moving and stimulating the muscles of the scapula raises one\textquotesingle s basal metabolism. }

\par{106. 世界中の ${\overset{\textnormal{きょうさんしゅぎしゃ}}{\text{共産主義者}}}$ はロシアや中国などといった国々に ${\overset{\textnormal{げんわく}}{\text{幻惑}}}$ されていた。 \hfill\break
The communists of the world had been enchanting by countries such as Russia and China. }

\par{107. 今回のケースでは、 ${\overset{\textnormal{ぼうこう}}{\text{暴行}}}$ が複数回に ${\overset{\textnormal{およ}}{\text{及}}}$ んでいたため、 ${\overset{\textnormal{きんこ}}{\text{禁固}}}$ 150年の ${\overset{\textnormal{はんけつ}}{\text{判決}}}$ となった模様だ。 \hfill\break
In this case, because the assaults had amounted to several times, it seems that the verdict was a 150 year prison sentence. }

\par{108. ${\overset{\textnormal{だいよんじゅうしちじょう}}{\text{第四十七条}}}$  ${\overset{\textnormal{きこう}}{\text{機構}}}$ は、 ${\overset{\textnormal{きゅうさいきゅうふ}}{\text{救済給付}}}$ の支給に要する費用に ${\overset{\textnormal{あ}}{\text{充}}}$ てるため、 ${\overset{\textnormal{いしわた}}{\text{石綿}}}$ の ${\overset{\textnormal{しようりょう}}{\text{使用量}}}$ 、 ${\overset{\textnormal{していしっぺい}}{\text{指定疾病}}}$ の ${\overset{\textnormal{はっせい}}{\text{発生}}}$ の ${\overset{\textnormal{じょうきょう}}{\text{状況}}}$ その ${\overset{\textnormal{た}}{\text{他}}}$ の事情を ${\overset{\textnormal{かんあん}}{\text{勘案}}}$ して ${\overset{\textnormal{せいれい}}{\text{政令}}}$ で定める要件に該当する ${\overset{\textnormal{じぎょうぬし}}{\text{事業主}}}$ から、 ${\overset{\textnormal{まいねんど}}{\text{毎年度}}}$ 、 ${\overset{\textnormal{とくべつきょしゅつきん}}{\text{特別拠出金}}}$ を ${\overset{\textnormal{ちょうしゅう}}{\text{徴収}}}$ する。 \hfill\break
From the 石綿による健康被害の救済に関する法律 \hfill\break
Article 47: The Agency will impose special contributions annually to business owners who meet the requirements prescribed by the ordinance—taking into consideration the usage amount of asbestos, outbreak conditions of designated diseases, and other circumstances—in order to allot the costs for providing relief benefits. }
 
\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Purple spot & しはん & 紫斑 & 紫班 & Acclimation & じゅんか & 馴化 & 順化 \\ \cline{1-8}

Compliance & じゅんしゅ & 遵守 & 順守 & Law observance & じゅんぽう & 遵法 & 順法 \\ \cline{1-8}

Homeliness & じゅんぼく & 醇朴 & 純朴 & Swift horse & しゅんば & 駿馬 & 俊馬 \\ \cline{1-8}

Loquacity & じょうぜつ & 饒舌 & 冗舌 & Tapeworm & じょうちゅう & 絛虫 & 条虫 \\ \cline{1-8}

Meal & しょくじ & 食餌 & 食事 & Newly compiled & しんせん & 新撰 & 新選 \\ \cline{1-8}

Heart rate & しんぱく & 心搏 & 心拍 & Compiler & せんじゃ & 撰者 & 選者 \hfill\break
\\ \cline{1-8}

Anthology & せんしゅう & 撰集 & 選集 & Compiling & せんしゅう & 撰修 & 選修 \\ \cline{1-8}

Vanguard & せんぺい & 尖兵 & 先兵 & Prominence & たいとう & 擡頭 & 台頭 \\ \cline{1-8}

Unattainable goal & たかねのはな & 高嶺の花 & 高根の花 & Determined\slash resolute & だんこ & 断乎 & 断固 \\ \cline{1-8}

Lottery\slash raffle & ちゅうせん & 抽籤 & 抽選 & Bird's-eye view & ちょうかん & 鳥瞰 & 鳥観 \\ \cline{1-8}

\end{ltabulary}
 
\par{109. 10歳の娘がアレルギー ${\overset{\textnormal{せい}}{\text{性}}}$ ${\overset{\textnormal{しはんびょう}}{\text{紫斑病}}}$ を ${\overset{\textnormal{はっしょう}}{\text{発症}}}$ しました。 \hfill\break
My ten year old daughter has developed an allergic purpura. }

\par{110. きょうは、前回と異なる種類のキュウリも同じ環境に ${\overset{\textnormal{じゅんか}}{\text{順化}}}$ するか検証してみたが、驚くことに ${\overset{\textnormal{にさんかたんそのうど}}{\text{二酸化炭素濃度}}}$ の ${\overset{\textnormal{ていか}}{\text{低下}}}$ が ${\overset{\textnormal{みしょう}}{\text{実生}}}$ の ${\overset{\textnormal{こうごうせい}}{\text{光合成}}}$ に ${\overset{\textnormal{あくえいきょう}}{\text{悪影響}}}$ を ${\overset{\textnormal{およ}}{\text{及}}}$ ぼしてしまって、成長を妨げたことがわかりました。 \hfill\break
Today, we tested to see whether a different species of cucumber would accumulate to the same environment as the last one, but to our surprise, we found out that the decrease in the density of carbon dioxide had adverse effects on the photosynthesis of the seedlings, impairing its growth. }

\par{111. こういう事件が今でも多発していることに ${\overset{\textnormal{かんが}}{\text{鑑}}}$ みると、純朴すぎて利用されまくるという人が世間には多いことに ${\overset{\textnormal{きぐ}}{\text{危惧}}}$ を覚えるでしょう。 \hfill\break
In the light of this sort of incident occurring frequently even now, you get apprehended with the sense of uneasiness that there is a lot of people in this world that are too naïve and constantly used. }

\par{112. ${\overset{\textnormal{じょうちゅう}}{\text{条虫}}}$ を ${\overset{\textnormal{たいじ}}{\text{退治}}}$ するとともに ${\overset{\textnormal{たいがい}}{\text{体外}}}$ に ${\overset{\textnormal{はいしゅつ}}{\text{排出}}}$ するための ${\overset{\textnormal{ちりょう}}{\text{治療}}}$ を行う必要があります。 \hfill\break
Along with exterminating the tapeworms, treatment for expelling them from the body is necessary. }

\par{113. ${\overset{\textnormal{しんぱくすう}}{\text{心拍数}}}$ を測るためには、心拍計を使えばわかるはずです。 \hfill\break
To measure heart rate, you should be able to figure it out by using a heart rate monitor. }

\par{114. ${\overset{\textnormal{ちょうかんず}}{\text{鳥瞰図}}}$ を英語に訳すと何と言うんでしょうか。 \hfill\break
How would you translate “bird\textquotesingle s-eye view” into English? }

\par{115. もう一度 ${\overset{\textnormal{ちゅうせん}}{\text{抽籤}}}$ してみたが、また落ちてしまった。 \hfill\break
I cast a lot once more, but I lost again. }

\par{116. きのう、 ${\overset{\textnormal{いしかわじゅん}}{\text{石川淳}}}$ という作家の ${\overset{\textnormal{しんしょばんせんじゅう}}{\text{新書版撰集}}}$ を手に入れました。 \hfill\break
I obtained a selection of new published works of the author Jun Ishikawa. }

\par{117. 私もパク・クネ政権に ${\overset{\textnormal{だんこ}}{\text{断固}}}$ として反対です。 \hfill\break
I too am adamantly against Park Geun Hye\textquotesingle s administration. }

\par{118. 中国政府も他の国と同じように、何故トランプ氏がこんなに強く ${\overset{\textnormal{たいとう}}{\text{台頭}}}$ してきたのかすっかり ${\overset{\textnormal{こんわく}}{\text{困惑}}}$ しているに違いない。 \hfill\break
The Chinese government, in the same way as other nations, is also no doubt utterly puzzled as to why Mr. Trump has so strongly come to the forefront. }

\par{119. いつも ${\overset{\textnormal{じょうぜつ}}{\text{饒舌}}}$ な私が思わず無言になってしまった。 \hfill\break
I myself, who is always verbose, went silent. }

\par{120. 昔ながらの ${\overset{\textnormal{でんとう}}{\text{伝統}}}$ を ${\overset{\textnormal{じゅんしゅ}}{\text{遵守}}}$ しながらも、現代でしか味わえない感覚をも ${\overset{\textnormal{ついきゅう}}{\text{追求}}}$ し、人生をより ${\overset{\textnormal{しきさいゆた}}{\text{色彩豊}}}$ かにすることができる。 \hfill\break
One is able to have a richer colorful life by also pursuing sensations only possible to experience in the present whilst acting in accordance to time-honored traditions. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Tossing and turning in bed & てんてんはんそく & 輾転反側 & 展転反側 & Buttocks & でんぶ & 臀部 & 殿部 \\ \cline{1-8}

Winning (lottery) & とうせん & 当籤 & 当選 \hfill\break
& Ruins & はいきょ & 廃墟 & 廃虚 \\ \cline{1-8}

Comprador & ばいべん & 買辦 & 買弁 & Chalk & はくあ & 白堊 & 白亜 \\ \cline{1-8}

Pulsation & はくどう & 搏動 & 拍動 & Disturbance & はらん & 波瀾 & 波乱 \\ \cline{1-8}

Speck\slash fleck & はんてん & 斑点 & 班点 & Incompetence & ひさい & 菲才 & 非才 \hfill\break
\\ \cline{1-8}

Aimlessly & ひょうぜん & 飄然 & 漂然 & Revealing & ひれき & 披瀝 & 披歴 \\ \cline{1-8}

Allegory & ふうゆ & 諷喩 & 風諭 & Expatiation & ふえん & 敷衍 & 敷延 \\ \cline{1-8}

Fractional distillation & ぶんりゅう & 分溜 & 分留 & Pigtail\slash queue & べんぱつ & 辮髪 & 弁髪 \\ \cline{1-8}

Flatness & へんぺい & 扁平 & 偏平 & Prevention & ぼうあつ & 防遏 & 防圧 \hfill\break
\\ \cline{1-8}

Bearing\slash presenting & ほうじ & 捧持 & 奉持 & Mellow & ほうじゅん & 芳醇 & 芳純 \\ \cline{1-8}

\end{ltabulary}

\par{121. ${\overset{\textnormal{きしょうじかん}}{\text{起床時間}}}$ の5時半まで ${\overset{\textnormal{てんてんはんそく}}{\text{輾転反側}}}$ していました。 \hfill\break
I was tossing and turning in bed until 5:30, the time I get up. }

\par{122. ${\overset{\textnormal{でんぶ}}{\text{臀部}}}$ の痛みは ${\overset{\textnormal{ろうにゃくなんにょ}}{\text{老若男女}}}$ ${\overset{\textnormal{と}}{\text{問}}}$ わず ${\overset{\textnormal{あらわ}}{\text{現}}}$ れる ${\overset{\textnormal{しょうじょう}}{\text{症状}}}$ だ。 \hfill\break
Gluteal pain is a symptom that emerges without regard to age or sex. }

\par{123. ${\overset{\textnormal{はいきょ}}{\text{廃墟}}}$ が ${\overset{\textnormal{くず}}{\text{崩}}}$ れ始めたところ、 ${\overset{\textnormal{まほう}}{\text{魔法}}}$ の ${\overset{\textnormal{じゅうたん}}{\text{絨毯}}}$ が来てくれた。 \hfill\break
Just when the ruins began to crumble, a magic carpet came to the rescue. }

\par{124. ${\overset{\textnormal{たからくじ}}{\text{宝籤}}}$ に ${\overset{\textnormal{とうせん}}{\text{当籤}}}$ したら ${\overset{\textnormal{よくとし}}{\text{翌年}}}$ の ${\overset{\textnormal{ぜいきん}}{\text{税金}}}$ はどうなるんだろう。 \hfill\break
I wonder what my taxes would be like the next year after winning the lottery. }

\par{125. ${\overset{\textnormal{はいかい}}{\text{俳諧}}}$ は、 ${\overset{\textnormal{きほんてき}}{\text{基本的}}}$ には社会に対して ${\overset{\textnormal{ひていてき}}{\text{否定的}}}$ な ${\overset{\textnormal{ふうゆ}}{\text{風諭}}}$ を ${\overset{\textnormal{ふく}}{\text{含}}}$ まない。 \hfill\break
Haikai poems, as a rule, do not contain negative allegories toward society. }

\par{126. ${\overset{\textnormal{べんぱつ}}{\text{辮髪}}}$ は ${\overset{\textnormal{だんし}}{\text{男子}}}$ が ${\overset{\textnormal{とうはつ}}{\text{頭髪}}}$ を ${\overset{\textnormal{そ}}{\text{剃}}}$ り、 ${\overset{\textnormal{こうとうぶ}}{\text{後頭部}}}$ だけを長く ${\overset{\textnormal{の}}{\text{伸}}}$ ばして ${\overset{\textnormal{あ}}{\text{編}}}$ み、 ${\overset{\textnormal{はいご}}{\text{背後}}}$ に長く ${\overset{\textnormal{た}}{\text{垂}}}$ らす ${\overset{\textnormal{かみがた}}{\text{髪型}}}$ のことである。 \hfill\break
The queue is a male hairstyle with the head shaven, the hair at the back of the head lengthened and braided and hanging long down the back. }

\par{127. テロに対しての ${\overset{\textnormal{ぼうあつ}}{\text{防遏}}}$ は、人によって ${\overset{\textnormal{きず}}{\text{築}}}$ かれ、人によって守られる。 \hfill\break
Hindrance against terrorism is built by people and protected by people. }

\par{128. ${\overset{\textnormal{あま}}{\text{甘}}}$ くて ${\overset{\textnormal{ほうじゅん}}{\text{芳醇}}}$ なワインを楽しましょう。 \hfill\break
Enjoy sweet and mellow wine. }

\par{129. 一週間くらい前に ${\overset{\textnormal{へんぺい}}{\text{扁平}}}$ なイボができ、色は ${\overset{\textnormal{うす}}{\text{薄}}}$ い ${\overset{\textnormal{かっしょく}}{\text{褐色}}}$ で、あまり目立ちませんが、ちょっと痛いのでどうしたらいいか分かりません。 \hfill\break
About a week ago, a flat wart formed; the color is a light brown and it doesn\textquotesingle t really stand out, but since it does hurt a little, I\textquotesingle m not sure what to do. }

\par{130. ${\overset{\textnormal{じゅうじか}}{\text{十字架}}}$ を ${\overset{\textnormal{ほうじ}}{\text{捧持}}}$ する人を「クロスベアラー」と呼ぶ。 \hfill\break
A person holding a crucifix is called a “cross-bearer.” }

\par{131. ${\overset{\textnormal{はくあき}}{\text{白亜紀}}}$ を ${\overset{\textnormal{だいひょう}}{\text{代表}}}$ する ${\overset{\textnormal{きょうりゅう}}{\text{恐竜}}}$ といえば、T・レックスでしょう。 \hfill\break
Speaking of a dinosaur that represents the Cretaceous Period, there\textquotesingle s T Rex. }

\par{132. ${\overset{\textnormal{つうかきき}}{\text{通貨危機}}}$ が ${\overset{\textnormal{はらん}}{\text{波乱}}}$ を ${\overset{\textnormal{ま}}{\text{巻}}}$ き ${\overset{\textnormal{お}}{\text{起}}}$ こしている。 \hfill\break
The monetary crisis is causing a storm of troubles. }

\par{133. ${\overset{\textnormal{せんがく}}{\text{浅学}}}$ にして ${\overset{\textnormal{ひさい}}{\text{菲才}}}$ の ${\overset{\textnormal{み}}{\text{身}}}$ 。 \hfill\break
Lack of learning is a lack of ability. }

\par{134. インドを ${\overset{\textnormal{とうち}}{\text{統治}}}$ するイギリス人の ${\overset{\textnormal{せかいかん}}{\text{世界観}}}$ を ${\overset{\textnormal{ひれき}}{\text{披瀝}}}$ する映画である。 \hfill\break
This is a movie which reveals the worldview of Englishmen who ruled India. }

\par{135. 下記が ${\overset{\textnormal{かれい}}{\text{加齢}}}$ による ${\overset{\textnormal{はくどう}}{\text{拍動}}}$ の変化を示す ${\overset{\textnormal{ずひょう}}{\text{図表}}}$ である。 \hfill\break
Below is a chart which shows the change in heart beat due to aging. }

\par{136. 足に赤い ${\overset{\textnormal{はんてん}}{\text{斑点}}}$ ができた。 \hfill\break
A red speckle formed on my foot. }

\par{137. ${\overset{\textnormal{ひょうぜん}}{\text{飄然}}}$ として風の ${\overset{\textnormal{ごと}}{\text{如}}}$ く ${\overset{\textnormal{き}}{\text{消}}}$ え ${\overset{\textnormal{さ}}{\text{去}}}$ る。 \hfill\break
To fade away aimlessly like the wind. }

\par{138.これはキリスト教の基本的な立場を ${\overset{\textnormal{ふえん}}{\text{敷衍}}}$ したものと考えられます。 \hfill\break
This is thought to be an amplification of the fundamentals of Christianity. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Swelling & ぼうちょう & 膨脹 & 膨張 & Stronghold & ほうるい & 堡塁 & 保塁 \\ \cline{1-8}

Thumb & ぼし & 拇指 & 母指 & Pulse & みゃくはく & 脈搏 & 脈拍 \\ \cline{1-8}

Picturesque & めいび & 明媚 & 明美 & Fellowship & ゆうぎ & 友誼 & 友宜 \\ \cline{1-8}

Fine horse & ゆうしゅん & 優駿 & 優俊 & Calmly & ゆうぜん & 悠然 & 裕然 \\ \cline{1-8}

Transport ship & ゆそうせん & 油槽船 & 油送船 & Public opinion & よろん & 輿論 & 世論 \\ \cline{1-8}

Straggling & らくご & 落伍 & 落後 & Over-hunting & らんかく & 濫獲 & 乱獲 \\ \cline{1-8}

Over-mining & らんくつ & 濫掘 & 乱掘 & Over-production & らんさく & 濫作 & 乱作 \\ \cline{1-8}

Over-production & らんぞう & 濫造 & 乱造 & Desultory reading & らんどく & 濫読 & 乱読 \\ \cline{1-8}

Reckless deforestation & らんばつ & 濫伐 & 乱伐 & Reckless firing & らんぱつ & 濫発 & 乱発 \\ \cline{1-8}

Laving spending & らんぴ & 濫費 & 乱費 & Misappropriation & らんよう & 濫用 & 乱用 \\ \cline{1-8}

\end{ltabulary}
 
\par{139. 宇宙が ${\overset{\textnormal{ぼうちょう}}{\text{膨張}}}$ していることが理解されるとすぐに、宇宙は過去には今より小さかったはずであることが認識された。 \hfill\break
Immediately once it was understood that the universe was expanding, it was recognized that the universe in the past had to have been smaller than it is now. }

\par{140. 緊張した時など ${\overset{\textnormal{みゃくはく}}{\text{脈拍}}}$ が早くなることがある。 \hfill\break
There are times when one\textquotesingle s pulse gets faster such as when one is tense. }

\par{141. ${\overset{\textnormal{ぼしたいこうせい}}{\text{母指対向性}}}$ とは ${\overset{\textnormal{れいちょうるい}}{\text{霊長類}}}$ の ${\overset{\textnormal{とくちょう}}{\text{特徴}}}$ の一つである。 \hfill\break
The ability to oppose the thumb is one characteristic of primates. }

\par{142. どんな ${\overset{\textnormal{めいび}}{\text{明媚}}}$ な ${\overset{\textnormal{さんすい}}{\text{山水}}}$ でも近寄ってみれば ${\overset{\textnormal{こぎたな}}{\text{小汚}}}$ い ${\overset{\textnormal{ざっそう}}{\text{雑草}}}$ などがその ${\overset{\textnormal{したじ}}{\text{下地}}}$ になって ${\overset{\textnormal{すそ}}{\text{裾}}}$ を引いているのでしょう。 \hfill\break
No matter how scenic landscape may be, when you approach it, you will surely see that dirty weeds and such form the groundwork of the scenery and trail the bottom of it. }

\par{143. 社会から ${\overset{\textnormal{らくご}}{\text{落伍}}}$ していってしまう。 \hfill\break
To straggle out of society. }

\par{144. 信念をやり通せば世論を動かすことができる。 \hfill\break
When you achieve your convictions, you are able to mobilize public opinion. }

\par{145. その魚は ${\overset{\textnormal{らんかく}}{\text{乱獲}}}$ のため ${\overset{\textnormal{ぜつめつきぐしゅ}}{\text{絶滅危惧種}}}$ となっている。 \hfill\break
That fish has become an endangered species due to overfishing. }

\par{146. ${\overset{\textnormal{かいこけん}}{\text{解雇権}}}$ を ${\overset{\textnormal{らんよう}}{\text{濫用}}}$ すると、解雇は ${\overset{\textnormal{むこう}}{\text{無効}}}$ となる。 \hfill\break
If you abuse the right to discharge (an employee), the discharge becomes invalid. }

\par{147. ${\overset{\textnormal{しへい}}{\text{紙幣}}}$ の ${\overset{\textnormal{らんぱつ}}{\text{乱発}}}$ をするとなぜ ${\overset{\textnormal{ぶっか}}{\text{物価}}}$ は ${\overset{\textnormal{じょうしょう}}{\text{上昇}}}$ するのか。 \hfill\break
Why do prices increase when you over-issue bank notes? }

\par{148. 金持ちの家に生まれた子供は必ずしも ${\overset{\textnormal{ざいさん}}{\text{財産}}}$ を ${\overset{\textnormal{らんぴ}}{\text{乱費}}}$ するとは限らない。 \hfill\break
Children who are born in wealthy households do not always necessarily reckless spend their fortunes. }

\par{149. 内容のない小説を乱作する作家になんてなりたくない。 \hfill\break
I do not want to become the kind of author who mass-produces works which have no substance. }

\begin{ltabulary}{|P|P|P|P|P|P|P|P|}
\hline 

意味 & 読み & 元の用字 & 変化後 & 意味 & 読み & 元の用字 & 変化後 \\ \cline{1-8}

Guardrail & らんかん & 欄杆・闌干 & 欄干 & Water brash & りゅういん & 溜飲 & 留飲 \\ \cline{1-8}

Distillation & りゅうしゅつ & 溜出 & 留出 & Outdoing & りょうが & 凌駕 & 陵駕 \\ \cline{1-8}

Inclination & りょうけん & 料簡 & 了見 & Shouts of encouragement & れいせい & 厲声 & 励声 \\ \cline{1-8}

Two reams & れん \hfill\break
& 嗹 & 連 & Japanese person & わじん & 倭人 & 和人 \\ \cline{1-8}

\end{ltabulary}
 
\par{150. 彼は ${\overset{\textnormal{やけい}}{\text{夜景}}}$ に ${\overset{\textnormal{みと}}{\text{見惚}}}$ れて、 ${\overset{\textnormal{らんかん}}{\text{欄干}}}$ に ${\overset{\textnormal{もた}}{\text{凭}}}$ れて ${\overset{\textnormal{たたず}}{\text{佇}}}$ んだ。 \hfill\break
He stood and gazed in wonder at the nightscape while leaning against the bannister. }

\par{151. 真実の愛は全てを ${\overset{\textnormal{りょうが}}{\text{凌駕}}}$ する！ \hfill\break
True love is superior over everything. }

\par{152. ${\overset{\textnormal{わじん}}{\text{倭人}}}$ が単純に日本人になったと断言できない。 \hfill\break
One cannot affirmatively state that the “wajin” simply became the Japanese people. }

\par{153. ${\overset{\textnormal{かくせいひん}}{\text{各製品}}}$ の ${\overset{\textnormal{りゅうしゅつおんど}}{\text{留出温度}}}$ を ${\overset{\textnormal{すいてい}}{\text{推定}}}$ するモデルの開発を進めている。 \hfill\break
(They are) furthering development of a model to estimate the distillation temperature of each product. }

\par{154. 日本代表が勝利して ${\overset{\textnormal{りゅういん}}{\text{溜飲}}}$ が下がった。 \hfill\break
I felt relieved when the Japanese delegation was victorious. }

\par{155. 見知らぬ男に時間で買われるのを、こんな息苦しくなるような部屋で、雑誌を読みながら何時間でも待っているなんて、今時の娘たちは、どういう ${\overset{\textnormal{りょうけん}}{\text{了見}}}$ なのか、滝沢にはまったく理解できない。 \hfill\break
Takizawa could not understand at all what sort of motives female youth these days had that drove them to wait for hours as they read magazines in such a suffocating room as this to be bought by the hour by men they didn\textquotesingle t know. \hfill\break
From 凍える牙 by 乃南アサ. }
    