    
\chapter{交ぜ書き}

\begin{center}
\begin{Large}
第364課: 交ぜ書き 
\end{Large}
\end{center}
 
\par{ 交ぜ書き is when a single phrase consists of 漢字 that have been substituted with かな. This practice came about from language policies pushed by the National Language Council to curb the difficulty of learning Japanese orthography. }

\par{ Today, the list of characters for general use has recently been updated in light of the fact that typing has made such a list irrelevant, as well as other changes to the orthography including 交ぜ書き. Although substituting 漢字 for かな has always been at the discretion of the writer, strictly abiding by the 常用漢字表 has resulted in words commonly written half in 漢字 and half in かな in the media and official texts. }
      
\section{用例}
 
\par{ The number of words written in 交ぜ書き is rather high, but change in public opinion toward the abandoning of 交ぜ書き as much as possible is becoming more prevalent. Because the practice is also used in the gradual implementation of 漢字 in education*, this is a facet of the writing system that will have to be dealt with on a case-by-case basis. Nevertheless, below are real examples extracted from news articles to demonstrate this phenomenon. }

\par{*: Even this practice is being phased out by utilizing ふりがな over untaught characters. }

\par{1. 女性に多い ${\overset{\textnormal{しきゅう}}{\text{子宮}}}$ \textbf{がん }の一種、「 \textbf{子宮体がん }」の ${\overset{\textnormal{ちりょう}}{\text{治療}}}$ の ${\overset{\textnormal{さい}}{\text{際}}}$ 、がん ${\overset{\textnormal{さいぼう}}{\text{細胞}}}$ が残っているかどうかを ${\overset{\textnormal{がぞう}}{\text{画像}}}$ で ${\overset{\textnormal{しんだん}}{\text{診断}}}$ できる ${\overset{\textnormal{けんさほう}}{\text{検査法}}}$ を ${\overset{\textnormal{ふくいだいがく}}{\text{福井大学}}}$ の ${\overset{\textnormal{いりょう}}{\text{医療}}}$ チームが開発し、治療を終えた女性が今月、出産したと発表しました。 \hfill\break
The medical treatment team of the University of Fukui has developed a way to examine whether cancer cells remain with image diagnosis for “cancer of the uterine body,” a kind of uterine cancer prevalent in women. \hfill\break
From NHK on 1\slash 30\slash 17. }

\par{\textbf{漢字 Note }: 子宮がん ＝ 子宮癌. }

\par{2. おたふくかぜはムンプスウイルスが原因の ${\overset{\textnormal{かんせんしょう}}{\text{感染症}}}$ で、 ${\overset{\textnormal{おも}}{\text{主}}}$ にせきなどによる ${\overset{\textnormal{ひ}}{\text{飛}}}$ \textbf{まつ }で感染し、発熱や耳の下の ${\overset{\textnormal{は}}{\text{腫}}}$ れなどを引き起こします。 \hfill\break
The mumps is an infectious disease caused by the mumps virus. Infection is mainly spread by spraying from coughing, and it induces fever and swelling under the ears. \hfill\break
From NHK on 1\slash 29\slash 17. }

\par{\textbf{漢字 Note }: 飛まつ ＝ 飛沫. }

\par{3. そして、 ${\overset{\textnormal{いちれん}}{\text{一連}}}$ の ${\overset{\textnormal{じっけんちゅう}}{\text{実験中}}}$ に、 ${\overset{\textnormal{のう}}{\text{脳}}}$ の中でどのように記憶が作られたのか ${\overset{\textnormal{かんさつ}}{\text{観察}}}$ したところ、 ${\overset{\textnormal{でんきしげき}}{\text{電気刺激}}}$ の ${\overset{\textnormal{きょうふ}}{\text{恐怖}}}$ の記憶と ${\overset{\textnormal{あま}}{\text{甘}}}$ い水の記憶は、それぞれ脳の \textbf{へんとう }${\overset{\textnormal{たい}}{\text{}}}$ と呼ばれる場所で異なる ${\overset{\textnormal{しんけい}}{\text{神経}}}$ 細胞の ${\overset{\textnormal{しゅうだん}}{\text{集団}}}$ に ${\overset{\textnormal{たくわ}}{\text{蓄}}}$ えられましたが、２つの細胞の集団が、一部で重なり合うと、両方の記憶がつながることがわかったということです。 \hfill\break
And now during a series of experiments, after observing how memories are created inside the brain, memories of fear toward electric stimulus and memories of sweet water were each respectively collected in different neuron masses in places in the brain called amygdala, but it was fond out that when a portion the two cell masses overlap, both sets of memories are connected. \hfill\break
From NHK on 1\slash 29\slash 17. }

\par{\textbf{漢字 Note }: へんとう体 ＝ 扁桃体. }

\par{4. 専門家は、ウイルスの \textbf{まん }${\overset{\textnormal{えん}}{\text{延}}}$ が ${\overset{\textnormal{きわ}}{\text{極}}}$ めて ${\overset{\textnormal{しんこく}}{\text{深刻}}}$ なレベルにあると ${\overset{\textnormal{けいしょう}}{\text{警鐘}}}$ を ${\overset{\textnormal{な}}{\text{鳴}}}$ らしています。 \hfill\break
Experts are raising the alarm that the spread of the virus has reached an exceedingly severe level. \hfill\break
From NHK on 1\slash 26\slash 17. }

\par{\textbf{漢字 Note }: まん延 ＝ 蔓延. }

\par{5. グループでは、あらかじめ \textbf{すい }${\overset{\textnormal{ぞう}}{\text{臓}}}$ ができないよう ${\overset{\textnormal{いでんし}}{\text{遺伝子}}}$ を操作したラットの ${\overset{\textnormal{じゅせいらん}}{\text{受精卵}}}$ にマウスのｉＰＳ細胞を入れ、 ${\overset{\textnormal{たいない}}{\text{体内}}}$ にマウスのすい臓をもつラットを作り出すことに成功しました。 \hfill\break
In the group, success was had in creating rats which had mouse pancreases in their bodies by injecting in advance mouse iPS cells into fertilized rat eggs, which had the genes operating the pancreas not work \hfill\break
From NHK on 1\slash 26\slash 17. }

\par{\textbf{漢字 Note }: すい臓 ＝ 膵臓. }

\par{6. 事故を受けて日本バレーボール ${\overset{\textnormal{きょうかい}}{\text{協会}}}$ は先月、 ${\overset{\textnormal{なかがいちじきかんとく}}{\text{中垣内次期監督}}}$ を \textbf{けん }${\overset{\textnormal{せきしょぶん}}{\text{責処分}}}$ にしています。 \hfill\break
In response to the accident, the Japan Volleyball Association gave an official reprimand Coach Nakagai for next term last month. \hfill\break
From NHK on 1\slash 30\slash 17. }

\par{\textbf{漢字 Note }: けん責 ＝ 譴責. }

\par{7. ${\overset{\textnormal{わかやまけんごぼうし}}{\text{和歌山県御坊市}}}$ の小中学校や ${\overset{\textnormal{ようちえん}}{\text{幼稚園}}}$ に ${\overset{\textnormal{かよ}}{\text{通}}}$ う子どもなど７００人以上が ${\overset{\textnormal{げり}}{\text{下痢}}}$ や \textbf{おう }${\overset{\textnormal{と}}{\text{吐}}}$ などの ${\overset{\textnormal{しょうじょう}}{\text{症状}}}$ \textbf{を }${\overset{\textnormal{うった}}{\text{訴}}}$ \textbf{えた問題で、 }${\overset{\textnormal{じどう}}{\text{児童}}}$ \textbf{や生徒の }${\overset{\textnormal{べん}}{\text{便}}}$ \textbf{からノロウイルスが }${\overset{\textnormal{けんしゅつ}}{\text{検出}}}$ \textbf{されたことが県の調査でわかりました。 \hfill\break
}In the case of over 700 children attending elementary schools and kindergartens of Gobo City in Wakayama City having symptoms including diarrhea and vomiting, it has been uncovered in a prefectural investigation that the norovirus has been detected in the stool of children and students. \hfill\break
From NHK on 1\slash 28\slash 17. }

\par{\textbf{漢字 Note }: おう吐 ＝ 嘔吐. }

\par{8. ２７日の ${\overset{\textnormal{はんけつ}}{\text{判決}}}$ で ${\overset{\textnormal{さっぽろちほうさいばんしょ}}{\text{札幌地方裁判所}}}$ の ${\overset{\textnormal{たじりかつみさいばんちょう}}{\text{田尻克已裁判長}}}$ は、「 ${\overset{\textnormal{きょうき}}{\text{凶器}}}$ を準備して ${\overset{\textnormal{おそ}}{\text{襲}}}$ ったうえ ${\overset{\textnormal{ひがいしゃ}}{\text{被害者}}}$ が倒れても ${\overset{\textnormal{しつ}}{\text{執}}}$ \textbf{よう }に ${\overset{\textnormal{ぼうこう}}{\text{暴行}}}$ を ${\overset{\textnormal{くわ}}{\text{加}}}$ え、 ${\overset{\textnormal{ざんぎゃく}}{\text{残虐}}}$ で ${\overset{\textnormal{けいかくてき}}{\text{計画的}}}$ な ${\overset{\textnormal{はんこう}}{\text{犯行}}}$ だ」と ${\overset{\textnormal{してき}}{\text{指摘}}}$ しました。 \hfill\break
In a verdict on the 27th, Judge Katsumi Tajiri of Sapporo District Court pointed out, “on top of having prepared a deadly weapon and assaulted the victim, further acts of relentless violence were dealt even after the victim had collapsed, which is a heartless and deliberate criminal offense.” \hfill\break
From NHK on 1\slash 27\slash 17. }

\par{\textbf{漢字 Note }: 執よう ＝ 執拗. }

\par{9. \textbf{うつ }${\overset{\textnormal{びょう}}{\text{病}}}$ を ${\overset{\textnormal{はっしょう}}{\text{発症}}}$ した ${\overset{\textnormal{もとしゃいん}}{\text{元社員}}}$ の男性に ${\overset{\textnormal{いほう}}{\text{違法}}}$ な ${\overset{\textnormal{ちょうじかんろうどう}}{\text{長時間労働}}}$ をさせたとして ${\overset{\textnormal{しょるいそうけん}}{\text{書類送検}}}$ された ${\overset{\textnormal{おおてでんき}}{\text{大手電機}}}$ メーカー、 ${\overset{\textnormal{みつびしでんき}}{\text{三菱電機}}}$ と当時の上司について、 ${\overset{\textnormal{よこはまちほうけんさつちょう}}{\text{横浜地方検察庁}}}$ は ${\overset{\textnormal{けんぎふじゅうぶん}}{\text{嫌疑不十分}}}$ で ${\overset{\textnormal{ふきそ}}{\text{不起訴}}}$ にしました。 \hfill\break
In regards to a police report having been sent to Mitsubishi Electric and the boss at said time for making a former male employee work illegal long shifts which induced depression, the Yokohama District Public Prosecutors Office has decided not to prosecute due to lack of suspicion. \hfill\break
From NHK on 1\slash 27\slash 17. }

\par{\textbf{漢字 Note }: うつ病 ＝ 鬱病・欝病. }

\par{10. 会場には、 ${\overset{\textnormal{ぎせいしゃ}}{\text{犠牲者}}}$ １９人分のろうそくが並べられ、はじめに全員で ${\overset{\textnormal{もく}}{\text{黙}}}$ \textbf{とう }をささげ、 ${\overset{\textnormal{ついとう}}{\text{追悼}}}$ しました。 \hfill\break
At the venue, candles for the 19 victims were lined up, and all participants began by offering a silent prayer as they mourned. \hfill\break
From NHK on 1\slash 26\slash 17. }

\par{\textbf{漢字 Note }: 黙とう ＝ 黙祷. }

\par{11. また、採用試験とは別に、教育委員会の人事に介入し、教育長が依頼を断るとどう喝したなどと指摘されている点については、持ち込まれた人事上の要望や意見を口頭で伝えたことは認めたものの、人事について指示したり、 \textbf{どう喝 }したりした事実はないと主張しました。 \hfill\break
Additionally, apart from the employment examination affair, in regards to being remonstrated for interfering with the personnel affairs of the Board of Education and having blackmailed the committee whenever he Superintendent of Education denied are request among other things, although he has admitted to having orally expressed his wishes and opinions for human affairs referred to him, he insists that there is no truth in him having made any instructions regarding personal affairs or having blackmailed. }

\par{\textbf{漢字 Note }: どう喝 ＝ 恫喝. }

\par{12. この中で、民間議員は、世界の経済政策の不確実性によって、日本経済も大きな影響を受ける可能性があるとして、自由貿易体制の維持・強化に貢献するとともに、変動に動じない \textbf{強じん }な国内経済を構築するため、引き続き賃金の上昇や働き方改革に取り組むよう提言しました。 \hfill\break
Amidst this, congressmen in civilian occupations have recommended that effort needs to be made for continued increase in wages and revolutionizing ways of work to build a tough domestic economy that isn\textquotesingle t perturbed by fluctuations along with contributions to the maintenance and strengthening of free trade agreements, citing that there is a possibility of the Japanese economy being greatly affected by the uncertainty in the world\textquotesingle s economic policies. \hfill\break
From NHK on 1\slash 25\slash 17. }

\par{\textbf{漢字 Note }: 強じん ＝ 強靭. }

\par{13. イージス ${\overset{\textnormal{じゅんようかん}}{\text{巡洋艦}}}$ はタグボートに \textbf{えい }${\overset{\textnormal{こう}}{\text{}}}$ され、 ${\overset{\textnormal{きち}}{\text{基地}}}$ に ${\overset{\textnormal{もど}}{\text{戻}}}$ ったということで、 \textbf{けが ${\overset{\textnormal{にん}}{\text{人}}}$ }はいませんでした。 \hfill\break
The Aegis cruiser was towed by a tug boat and returned to the base; no one was injured. \hfill\break
From NHK on 2\slash 1\slash 17. }

\par{\textbf{漢字 Note }: えい航 ＝ 曳航; けが人 ＝ 怪我人. }
 
\begin{ltabulary}{|P|P|P|P|P|}
\hline 

\# & Mazegaki & True Spelling & Romanization & Definition \\ \cline{1-5}

1. & 危ぐ & 危惧 & Kigu & Misgiving; apprehension \\ \cline{1-5}

2. & 愛きょう & 愛嬌 & Aikyou & Amiability \\ \cline{1-5}

3. & えん恨 & 怨恨 & Enkon & Enmity; grudge \\ \cline{1-5}

4. & 喝さい & 喝采 & Kassai & Applause \\ \cline{1-5}

5. & 真ちゅう & 真鍮 & Shinchuu & Brass \\ \cline{1-5}

6. & けん責 & 譴責 & Kenseki & Censure; reprimand \\ \cline{1-5}

7. & しゅん工、しゅん功 & 竣工、竣功 & Shunkou & Completion \\ \cline{1-5}

8. & ち密 & 緻密 & Chimitsu & Elaborate \\ \cline{1-5}

9. & 双へき & 双璧 & Souheki & Matchless people; pair of bright jewels \\ \cline{1-5}

10. & 全ぼう & 全貌 & Zenbou & The whole story \\ \cline{1-5}

11. & 晩さん & 晩餐 & Bansan & Dinner \\ \cline{1-5}

12. & 分水れい & 分水嶺 & Bunsuirei & Watershed; divide \\ \cline{1-5}

13. & 復しゅう & 復讐 & Fukushuu & Revenge \\ \cline{1-5}

14. & 和ぼく & 和睦 & Waboku & Making peace with; bury the hatchet \\ \cline{1-5}

15. & ら致 & 拉致 & Rachi & Abduction \\ \cline{1-5}

16. & らく印 & 烙印 & Rakuin & Brand; stigmatization \\ \cline{1-5}

17. & ほう助 & 幇助 & Houjo & Abetment \\ \cline{1-5}

18. & 明りょうな & 明瞭な & Meiryou na & Clear; obvious; precise; articulate \\ \cline{1-5}

19. & 憂うつ & 憂鬱 & Yuu'utsu & Depression; melancholy; gloom; dismal \\ \cline{1-5}

20. & 片りん & 片鱗 & Henrin & Glimpse \\ \cline{1-5}

21. & れき死 & 轢死 & Rekishi & Getting killed by being run over by a train \\ \cline{1-5}

22. & 美ぼう & 美貌 & Bibou & Beauty; good looks \\ \cline{1-5}

23. & ばい菌 & 黴菌 & Baikin & Germ; bacteria; bug \\ \cline{1-5}

24. & ぜい弱な & 脆弱な & Zeijaku na & Weak; fragile; vulnerable \\ \cline{1-5}

25. & 終えん & 終焉 & Shuuen & End; demise \\ \cline{1-5}

26. & さく裂 & 炸裂 & Sakuretsu & Burst; explosion \\ \cline{1-5}

27. & 研さん & 研鑽 & Kensan & Study \\ \cline{1-5}

28. & 花こう岩 & 花崗岩 & Kakougan & Granite \\ \cline{1-5}

29. & 欺まん & 欺瞞 & Giman & Falsehood; fraud; deceit \\ \cline{1-5}

30. & 外とう & 外套 & Gaitou & (Over)coat \\ \cline{1-5}

31. & 警ら & 警邏 & Keira & Patrol \\ \cline{1-5}

32. & う回 & 迂回 & Ukai & Detour \\ \cline{1-5}

33. & 苦もん & 苦悶 & Kumon & Agony; pang; anguish \\ \cline{1-5}

34. & かく乱 & 霍乱 & Kakuran & Heatstroke \\ \cline{1-5}

35. & 円すい & 円錐 & Ensui & Cone \\ \cline{1-5}

36. & 祈とう & 祈祷 & Kitou & Prayer; devotion \\ \cline{1-5}

37. & 啓もう & 啓蒙 & Keimou & Enlightenment \\ \cline{1-5}

38. & 進ちょく & 進捗 & Shinchoku & Progress \\ \cline{1-5}

39. & 断がい & 断崖 & Dangai & Cliff; bluff \\ \cline{1-5}

40. & どう喝 & 恫喝 & Doukatsu & Intimidation; threat \\ \cline{1-5}

41. & どう猛な & 獰猛な & Doumou na & Fierce; savage; ferocious \\ \cline{1-5}

42. & はえ縄 & 延縄 & Haenawa & Trawl line \\ \cline{1-5}

43. & 比ゆ & 比喩 & Hiyu & Similitude; comparison; metaphor; simile \\ \cline{1-5}

44. & らつ腕 & 辣腕 & Ratsuwan & Competent; shrewd \\ \cline{1-5}

45. & まい進 & 邁進 & Maishin & Pushing forward; strive \\ \cline{1-5}

46. & ろう城 & 篭城 & Roujou & Holding up a castle; shutting oneself up in a room \\ \cline{1-5}

47. & 払しょく & 払拭 & Fusshoku & Sweeping away; eradicate \\ \cline{1-5}

48. & はく製 & 剥製 & Hakusei & Stuffed specimen \\ \cline{1-5}

49. & 便せん & 便箋 & Binsen & Writing paper \\ \cline{1-5}

50. & 明せきな & 明晰な & Meiseki na & Clear; precise \\ \cline{1-5}

51. & 補てん & 補填 & Hoten & Compensation \\ \cline{1-5}

52. & 山ろく & 山麓 & Sanroku & The foot of a mountain \\ \cline{1-5}

53. & 混とん & 混沌 & Konton & Chaos \\ \cline{1-5}

54. & こん包 & 梱包 & Konpou & Packing \\ \cline{1-5}

55. & ごう音 & 轟音 & Gou'on & Roar \\ \cline{1-5}

56. & 完ぺきな & 完璧な & Kanpeki na & Perfect; faultless; impeccable \\ \cline{1-5}

57. & し好 & 嗜好 & Shikou & Taste; liking; fancy \\ \cline{1-5}

58. & そう快な & 爽快な & Soukai na & Refreshing \\ \cline{1-5}

59. & 投かん & 投函 & Toukan & Dispatch \\ \cline{1-5}

60. & 石こう & 石膏 & Sekkou & Gypsum; plaster; alabaster \\ \cline{1-5}

61. & 痴ほう & 痴呆 & Chihou & Dementia \\ \cline{1-5}

62. & 信ぴょう & 信憑 & Shinpyou & Credence \\ \cline{1-5}

63. & 同せい & 同棲 & Dousei & Living together \\ \cline{1-5}

64. & 棟りょう & 棟梁 & Touryou & Master builder \\ \cline{1-5}

65. & どん欲な & 貪欲な & Don'yoku na & Greedy; insatiable; voracious; mercenary; lustful; acquisitive \\ \cline{1-5}

66. & すい星 & 彗星 & Suisei & Comet \\ \cline{1-5}

67. & 軟こう & 軟膏 & Nankou & Ointment \\ \cline{1-5}

68. & ねつ造 & 捏造 & Netsuzou & Fabrication; forgery; fiction \\ \cline{1-5}

69. & 怒とう & 怒涛 & Dotou & Angry waves \\ \cline{1-5}

70. & 惨たん & 惨憺、惨澹 & Santan & Miserable; wretched \\ \cline{1-5}

71. & 破たん & 破綻 & Hatan & Bankruptcy \\ \cline{1-5}

72. & えい航 & 曳航 & Eikou & Tow \\ \cline{1-5}

73. & いん石 & 隕石 & Inseki & Meteorite \\ \cline{1-5}

74. & きゅう覚 & 嗅覚 & Kyuukaku & Smell; scent \\ \cline{1-5}

75. & 改しゅん、悔しゅん & 改悛、悔悛 & Kaishun & Penance \\ \cline{1-5}

76. & こう配 & 勾配 & Koubai & Inclination; slope; incline; grade; gradient \\ \cline{1-5}

77. & 常とう & 常套 & Joutou & Conventionality; triteness \\ \cline{1-5}

78. & せん光 & 閃光 & Senkou & Flash; sparkle; glance \\ \cline{1-5}

79. & 鉄つい & 鉄槌、鉄鎚 & Tettsui & Punish severely \\ \cline{1-5}

80. & 石けん & 石鹸 & Sekken & Soap \\ \cline{1-5}

81. & 黙とう & 黙祷 & Mokutou & Silent prayer; tacit prayer \\ \cline{1-5}

82. & 余ろく & 余禄 & Yoroku & Extra profit \\ \cline{1-5}

83. & もん絶 & 悶絶 & Monzetsu & Faint in agony \\ \cline{1-5}

84. & 辺ぴな & 辺鄙な & Henpi na & Remote; out-of-the-way \\ \cline{1-5}

85. & ろう屋 & 牢屋 & Rouya & Prison \\ \cline{1-5}

86. & 憤まん & 憤懣 & Funman & Changrin; anger; resentment; indignation; irritation \\ \cline{1-5}

87. & 老かいな & 老獪な & Roukai na & Cunning; craft \\ \cline{1-5}

88. & べっ甲 & 鼈甲 & Bekkou & Tortoiseshell \\ \cline{1-5}

89. & ひん死 & 瀕死 & Hinshi & Dying; moribund \\ \cline{1-5}

90. & 配ぜん & 配膳 & Haizen & Setting a table \\ \cline{1-5}

91. & たく鉢 & 托鉢 & Takuhatsu & Monk's begging \\ \cline{1-5}

92. & しゃく熱 & 灼熱 & Shakunetsu & Tropical; sultry \\ \cline{1-5}

93. & 教べん & 教鞭 & Kyouben & Teaching \\ \cline{1-5}

94. & 死がい & 死骸 & Shigai & Corpse; carcass; body; bone; remain \\ \cline{1-5}

95. & 軽べつ & 軽蔑 & Keibetsu & Contempt; scorn; disdain; despise \\ \cline{1-5}

96. & 安ど & 安堵 & Ando & Relief \\ \cline{1-5}

97. & 隠ぺい & 隠蔽 & Inpei & Concealment \\ \cline{1-5}

98. & けん引 & 牽引 & Ken'in & Traction \\ \cline{1-5}

99. & し尿 & 屎尿 & Shinyou & Human waste \\ \cline{1-5}

100. & でき愛 & 溺愛 & Dekiai & Dotage; fondness \\ \cline{1-5}

101. & ゆう出 & 湧出 & Yuushutsu & Gush; pour; flow \\ \cline{1-5}

102. & 泡まつ & 泡沫 & Houmatsu & Bubble \\ \cline{1-5}

103. & 酒こう & 酒肴 & Shukou & Sake and side dish; food and drink \\ \cline{1-5}

104. & 位はい & 位牌 & Ihai & Buddhist memorial tablet \\ \cline{1-5}

105. & 漏えい & 漏洩 & Rouei & Leakage \\ \cline{1-5}

106. & こん睡 & 昏睡 & Konsui & Coma \\ \cline{1-5}

107. & うっ憤 & 鬱憤 & Uppun & Grudge; anger; resentment \\ \cline{1-5}

108. & 巣くつ & 巣窟 & Soukutsu & Den; nest \\ \cline{1-5}

109. & 自ちょう & 自嘲 & Jichou & Self derision \\ \cline{1-5}

110. & 灯ろう & 灯篭 & Tourou & Garden lantern \\ \cline{1-5}

111. & ふ頭 & 埠頭 & Futou & Pier \\ \cline{1-5}

112. & 要さい & 要塞 & Yousai & Fortress \\ \cline{1-5}

113. & 辛らつな & 辛辣な & Shinratsu na & Severe; bitter; sharp; acid; pointed; poignant; biting; cutting \\ \cline{1-5}

114. & 霊きゅう車 & 霊柩車 & Reikyuusha & Hearse \\ \cline{1-5}

115. & ろ過 & 濾過 & Roka & Percolation; filtration \\ \cline{1-5}

116. & 秘けつ & 秘訣 & Hiketsu & Secret \\ \cline{1-5}

117. & 親ぼく & 親睦 & Shinboku & Friendship \\ \cline{1-5}

118. & だ液 & 唾液 & Daeki & Saliva \\ \cline{1-5}

119. & 洞くつ & 洞窟 & Doukutsu & Cave; ghotto; cavern \\ \cline{1-5}

120. & ばい煙 & 煤煙 & Baien & Soot; smoke \\ \cline{1-5}

121. & 失そう & 失踪 & Shissou & Disappearance \\ \cline{1-5}

122. & 改ざん & 改竄 & Kaizan & Falsification; manipulation; violence \\ \cline{1-5}

123. & えん曲 & 婉曲 & Enkyoku & Euphemism; roundabout \\ \cline{1-5}

124. & けん制 & 牽制 & Kensei & Tow \\ \cline{1-5}

125. & なつ印 & 捺印 & Natsuin & Signet; stigmitization \\ \cline{1-5}

126. & 土のう & 土嚢 & Donou & Sandbag \\ \cline{1-5}

127. & 装てん & 装填 & Souten & Charge; load \\ \cline{1-5}

128. & 抜てき & 抜擢 & Batteki & Selection; choose \\ \cline{1-5}

129. & はく離 & 剥離 & Hakuri & Detachment; coming\slash peeling off; separation \\ \cline{1-5}

130. & はく落 & 剥落 & Hakuraku & To peel off \\ \cline{1-5}

131. & ばく進 & 驀進 & Bakushin & Dash; hurtle \\ \cline{1-5}

132. & ねん挫 & 捻挫 & Nenza & Sprain; twist \\ \cline{1-5}

133. & しゅう恥 & 羞恥 & Shuuchi & Bashfulness; shyness \\ \cline{1-5}

134. & 遮へい & 遮蔽 & Shahei & Shield from \\ \cline{1-5}

135. & 付せん & 付箋 & Fusen & Post-it; tag; docket \\ \cline{1-5}

136. & 容ぼう & 容貌 & Youbou & Looks; appearance; feature; countenance \\ \cline{1-5}

137. & わい曲 & 歪曲 & Waikyoku & Distortion; twists and turns; twist \\ \cline{1-5}

138. & 落いん & 落胤 & Rakuin & The son\slash daughter of (literary); nobleman's illegitimate child \\ \cline{1-5}

139. & 防じん & 防塵 & Boujin & Dust protector \\ \cline{1-5}

140. & 排せつ & 排泄 & Haisetsu & Excrement \\ \cline{1-5}

141. & 殺りく & 殺戮 & Satsuriku & Slaughter; genocide; carnage \\ \cline{1-5}

142. & こん倒 & 昏倒 & Kontou & To faint unconscious \\ \cline{1-5}

143. & ばん回 & 挽回 & Bankai & Recovery; retrieve; restoration \\ \cline{1-5}

144. & 親せき & 親戚 & Shinseki & Relatives \\ \cline{1-5}

145. & 平たんな & 平坦な & Heitan na & Flat; level; even \\ \cline{1-5}

146. & 風ぼう & 風貌 & Fuubou & Feature; countenance \\ \cline{1-5}

147. & ひな形 & 雛形 & Hinagata & Model; form \\ \cline{1-5}

148. & ごう慢な & 傲慢な & Gouman na & Arrogant \\ \cline{1-5}

149. & こう着 & 膠着 & Kouchaku & Adhesion; agglutination; stalemate \\ \cline{1-5}

150. & かん腸 & 浣腸 & Kanchou & Enema \\ \cline{1-5}

151. & 炭そ菌 & 炭疽菌 & Tansokin & Anthrax \\ \cline{1-5}

152. & りょう線 & 稜線 & Ryousen & Mountain ridge \\ \cline{1-5}

153. & 悲そうな & 悲愴な & Hisou na & Sorrowful; mournful \\ \cline{1-5}

154. & えん世 & 厭世 & Ensei & Pessimism; misanthropy \\ \cline{1-5}

155. & がい骨 & 骸骨 & Gaikotsu & Skeleton \\ \cline{1-5}

156. & かい書 & 楷書 & Kaisho & Standard\slash block style \\ \cline{1-5}

157. & 刺しゅう & 刺繍 & Shishuu & Embroidery \\ \cline{1-5}

158. & けん怠 & 倦怠 & Kentai & Ennui; boredom \\ \cline{1-5}

159. & 卑わいな & 卑猥な & Hiwai na & Obscene; vulgar; profane \\ \cline{1-5}

160. & 標ぼう & 標榜 & Hyoubou & Profress \\ \cline{1-5}

161. & へん平 & 扁平 & Henpei & Flat; flat-footed; depressed \\ \cline{1-5}

162. & へき地 & 僻地 & Hekichi & Remote place; out-of-the-way place; isolated district; backwater \\ \cline{1-5}

163. & 流ちょうな & 流暢な & Ryuuchou na & Fluent; facile; smooth \\ \cline{1-5}

164. & きゅう舎 & 厩舎 & Kyuusha & Riding stable \\ \cline{1-5}

165. & 編さん & 編纂 & Hensan & Proofreading; editing; compilation; cutting \\ \cline{1-5}

166. & 風さい & 風采 & Fuusai & Appearance; look; figure \\ \cline{1-5}

167. & こん身 & 渾身 & Konshin & With all one's might \\ \cline{1-5}

168. & 飛しょう体 & 飛翔体 & Hishoutai & Flying object \\ \cline{1-5}

169. & 残がい & 残骸 & Zangai & Wreckage; ruin; wreck \\ \cline{1-5}

170. & 敏しょうな & 敏捷な & Binshou na & Lithe; prompt; nimble; speedy \\ \cline{1-5}

171. & 集じん & 集塵 & Shuujin & Collecting dust \\ \cline{1-5}

172. & 変ぼう & 変貌 & Henbou & Transfiguration; transformation \\ \cline{1-5}

173. & さい銭 & 賽銭 & Saisen & Money offered at a Shinto shrine \\ \cline{1-5}

174. & かく乱 & 撹乱 & Kakuran & Disturbance; perturb \\ \cline{1-5}

175. & 謙そん & 謙遜 & Kenson & Humility; humbleness; modisty \\ \cline{1-5}

176. & 覚せい剤 & 覚醒剤 & Kakuseizai & Stimulant \\ \cline{1-5}

\end{ltabulary}

\begin{ltabulary}{|P|P|P|P|P|}
\hline 

\# & Mazegaki & True Spelling & Romanization & Definition \\ \cline{1-5}

1. & 危ぐ & 危惧 & Kigu & Misgiving; apprehension \\ \cline{1-5}

2. & 愛きょう & 愛嬌 & Aikyou & Amiability \\ \cline{1-5}

3. & えん恨 & 怨恨 & Enkon & Enmity; grudge \\ \cline{1-5}

4. & 喝さい & 喝采 & Kassai & Applause \\ \cline{1-5}

5. & 真ちゅう & 真鍮 & Shinchuu & Brass \\ \cline{1-5}

6. & けん責 & 譴責 & Kenseki & Censure; reprimand \\ \cline{1-5}

7. & しゅん工、しゅん功 & 竣工、竣功 & Shunkou & Completion \\ \cline{1-5}

8. & ち密 & 緻密 & Chimitsu & Elaborate \\ \cline{1-5}

9. & 双へき & 双璧 & Souheki & Matchless people; pair of bright jewels \\ \cline{1-5}

10. & 全ぼう & 全貌 & Zenbou & The whole story \\ \cline{1-5}

11. & 晩さん & 晩餐 & Bansan & Dinner \\ \cline{1-5}

12. & 分水れい & 分水嶺 & Bunsuirei & Watershed; divide \\ \cline{1-5}

13. & 復しゅう & 復讐 & Fukushuu & Revenge \\ \cline{1-5}

14. & 和ぼく & 和睦 & Waboku & Making peace with; bury the hatchet \\ \cline{1-5}

15. & ら致 & 拉致 & Rachi & Abduction \\ \cline{1-5}

16. & らく印 & 烙印 & Rakuin & Brand; stigmatization \\ \cline{1-5}

17. & ほう助 & 幇助 & Houjo & Abetment \\ \cline{1-5}

18. & 明りょうな & 明瞭な & Meiryou na & Clear; obvious; precise; articulate \\ \cline{1-5}

19. & 憂うつ & 憂鬱 & Yuu'utsu & Depression; melancholy; gloom; dismal \\ \cline{1-5}

20. & 片りん & 片鱗 & Henrin & Glimpse \\ \cline{1-5}

21. & れき死 & 轢死 & Rekishi & Getting killed by being run over by a train \\ \cline{1-5}

22. & 美ぼう & 美貌 & Bibou & Beauty; good looks \\ \cline{1-5}

23. & ばい菌 & 黴菌 & Baikin & Germ; bacteria; bug \\ \cline{1-5}

24. & ぜい弱な & 脆弱な & Zeijaku na & Weak; fragile; vulnerable \\ \cline{1-5}

25. & 終えん & 終焉 & Shuuen & End; demise \\ \cline{1-5}

26. & さく裂 & 炸裂 & Sakuretsu & Burst; explosion \\ \cline{1-5}

27. & 研さん & 研鑽 & Kensan & Study \\ \cline{1-5}

28. & 花こう岩 & 花崗岩 & Kakougan & Granite \\ \cline{1-5}

29. & 欺まん & 欺瞞 & Giman & Falsehood; fraud; deceit \\ \cline{1-5}

30. & 外とう & 外套 & Gaitou & (Over)coat \\ \cline{1-5}

31. & 警ら & 警邏 & Keira & Patrol \\ \cline{1-5}

32. & う回 & 迂回 & Ukai & Detour \\ \cline{1-5}

33. & 苦もん & 苦悶 & Kumon & Agony; pang; anguish \\ \cline{1-5}

34. & かく乱 & 霍乱 & Kakuran & Heatstroke \\ \cline{1-5}

35. & 円すい & 円錐 & Ensui & Cone \\ \cline{1-5}

36. & 祈とう & 祈祷 & Kitou & Prayer; devotion \\ \cline{1-5}

37. & 啓もう & 啓蒙 & Keimou & Enlightenment \\ \cline{1-5}

38. & 進ちょく & 進捗 & Shinchoku & Progress \\ \cline{1-5}

39. & 断がい & 断崖 & Dangai & Cliff; bluff \\ \cline{1-5}

40. & どう喝 & 恫喝 & Doukatsu & Intimidation; threat \\ \cline{1-5}

41. & どう猛な & 獰猛な & Doumou na & Fierce; savage; ferocious \\ \cline{1-5}

42. & はえ縄 & 延縄 & Haenawa & Trawl line \\ \cline{1-5}

43. & 比ゆ & 比喩 & Hiyu & Similitude; comparison; metaphor; simile \\ \cline{1-5}

44. & らつ腕 & 辣腕 & Ratsuwan & Competent; shrewd \\ \cline{1-5}

45. & まい進 & 邁進 & Maishin & Pushing forward; strive \\ \cline{1-5}

46. & ろう城 & 篭城 & Roujou & Holding up a castle; shutting oneself up in a room \\ \cline{1-5}

47. & 払しょく & 払拭 & Fusshoku & Sweeping away; eradicate \\ \cline{1-5}

48. & はく製 & 剥製 & Hakusei & Stuffed specimen \\ \cline{1-5}

49. & 便せん & 便箋 & Binsen & Writing paper \\ \cline{1-5}

50. & 明せきな & 明晰な & Meiseki na & Clear; precise \\ \cline{1-5}

51. & 補てん & 補填 & Hoten & Compensation \\ \cline{1-5}

52. & 山ろく & 山麓 & Sanroku & The foot of a mountain \\ \cline{1-5}

53. & 混とん & 混沌 & Konton & Chaos \\ \cline{1-5}

54. & こん包 & 梱包 & Konpou & Packing \\ \cline{1-5}

55. & ごう音 & 轟音 & Gou'on & Roar \\ \cline{1-5}

56. & 完ぺきな & 完璧な & Kanpeki na & Perfect; faultless; impeccable \\ \cline{1-5}

57. & し好 & 嗜好 & Shikou & Taste; liking; fancy \\ \cline{1-5}

58. & そう快な & 爽快な & Soukai na & Refreshing \\ \cline{1-5}

59. & 投かん & 投函 & Toukan & Dispatch \\ \cline{1-5}

60. & 石こう & 石膏 & Sekkou & Gypsum; plaster; alabaster \\ \cline{1-5}

61. & 痴ほう & 痴呆 & Chihou & Dementia \\ \cline{1-5}

62. & 信ぴょう & 信憑 & Shinpyou & Credence \\ \cline{1-5}

63. & 同せい & 同棲 & Dousei & Living together \\ \cline{1-5}

64. & 棟りょう & 棟梁 & Touryou & Master builder \\ \cline{1-5}

65. & どん欲な & 貪欲な & Don'yoku na & Greedy; insatiable; voracious; mercenary; lustful; acquisitive \\ \cline{1-5}

66. & すい星 & 彗星 & Suisei & Comet \\ \cline{1-5}

67. & 軟こう & 軟膏 & Nankou & Ointment \\ \cline{1-5}

68. & ねつ造 & 捏造 & Netsuzou & Fabrication; forgery; fiction \\ \cline{1-5}

69. & 怒とう & 怒涛 & Dotou & Angry waves \\ \cline{1-5}

70. & 惨たん & 惨憺、惨澹 & Santan & Miserable; wretched \\ \cline{1-5}

71. & 破たん & 破綻 & Hatan & Bankruptcy \\ \cline{1-5}

72. & えい航 & 曳航 & Eikou & Tow \\ \cline{1-5}

73. & いん石 & 隕石 & Inseki & Meteorite \\ \cline{1-5}

74. & きゅう覚 & 嗅覚 & Kyuukaku & Smell; scent \\ \cline{1-5}

75. & 改しゅん、悔しゅん & 改悛、悔悛 & Kaishun & Penance \\ \cline{1-5}

76. & こう配 & 勾配 & Koubai & Inclination; slope; incline; grade; gradient \\ \cline{1-5}

77. & 常とう & 常套 & Joutou & Conventionality; triteness \\ \cline{1-5}

78. & せん光 & 閃光 & Senkou & Flash; sparkle; glance \\ \cline{1-5}

79. & 鉄つい & 鉄槌、鉄鎚 & Tettsui & Punish severely \\ \cline{1-5}

80. & 石けん & 石鹸 & Sekken & Soap \\ \cline{1-5}

81. & 黙とう & 黙祷 & Mokutou & Silent prayer; tacit prayer \\ \cline{1-5}

82. & 余ろく & 余禄 & Yoroku & Extra profit \\ \cline{1-5}

83. & もん絶 & 悶絶 & Monzetsu & Faint in agony \\ \cline{1-5}

84. & 辺ぴな & 辺鄙な & Henpi na & Remote; out-of-the-way \\ \cline{1-5}

85. & ろう屋 & 牢屋 & Rouya & Prison \\ \cline{1-5}

86. & 憤まん & 憤懣 & Funman & Changrin; anger; resentment; indignation; irritation \\ \cline{1-5}

87. & 老かいな & 老獪な & Roukai na & Cunning; craft \\ \cline{1-5}

88. & べっ甲 & 鼈甲 & Bekkou & Tortoiseshell \\ \cline{1-5}

89. & ひん死 & 瀕死 & Hinshi & Dying; moribund \\ \cline{1-5}

90. & 配ぜん & 配膳 & Haizen & Setting a table \\ \cline{1-5}

91. & たく鉢 & 托鉢 & Takuhatsu & Monk's begging \\ \cline{1-5}

92. & しゃく熱 & 灼熱 & Shakunetsu & Tropical; sultry \\ \cline{1-5}

93. & 教べん & 教鞭 & Kyouben & Teaching \\ \cline{1-5}

94. & 死がい & 死骸 & Shigai & Corpse; carcass; body; bone; remain \\ \cline{1-5}

95. & 軽べつ & 軽蔑 & Keibetsu & Contempt; scorn; disdain; despise \\ \cline{1-5}

96. & 安ど & 安堵 & Ando & Relief \\ \cline{1-5}

97. & 隠ぺい & 隠蔽 & Inpei & Concealment \\ \cline{1-5}

98. & けん引 & 牽引 & Ken'in & Traction \\ \cline{1-5}

99. & し尿 & 屎尿 & Shinyou & Human waste \\ \cline{1-5}

100. & でき愛 & 溺愛 & Dekiai & Dotage; fondness \\ \cline{1-5}

101. & ゆう出 & 湧出 & Yuushutsu & Gush; pour; flow \\ \cline{1-5}

102. & 泡まつ & 泡沫 & Houmatsu & Bubble \\ \cline{1-5}

103. & 酒こう & 酒肴 & Shukou & Sake and side dish; food and drink \\ \cline{1-5}

104. & 位はい & 位牌 & Ihai & Buddhist memorial tablet \\ \cline{1-5}

105. & 漏えい & 漏洩 & Rouei & Leakage \\ \cline{1-5}

106. & こん睡 & 昏睡 & Konsui & Coma \\ \cline{1-5}

107. & うっ憤 & 鬱憤 & Uppun & Grudge; anger; resentment \\ \cline{1-5}

108. & 巣くつ & 巣窟 & Soukutsu & Den; nest \\ \cline{1-5}

109. & 自ちょう & 自嘲 & Jichou & Self derision \\ \cline{1-5}

110. & 灯ろう & 灯篭 & Tourou & Garden lantern \\ \cline{1-5}

111. & ふ頭 & 埠頭 & Futou & Pier \\ \cline{1-5}

112. & 要さい & 要塞 & Yousai & Fortress \\ \cline{1-5}

113. & 辛らつな & 辛辣な & Shinratsu na & Severe; bitter; sharp; acid; pointed; poignant; biting; cutting \\ \cline{1-5}

114. & 霊きゅう車 & 霊柩車 & Reikyuusha & Hearse \\ \cline{1-5}

115. & ろ過 & 濾過 & Roka & Percolation; filtration \\ \cline{1-5}

116. & 秘けつ & 秘訣 & Hiketsu & Secret \\ \cline{1-5}

117. & 親ぼく & 親睦 & Shinboku & Friendship \\ \cline{1-5}

118. & だ液 & 唾液 & Daeki & Saliva \\ \cline{1-5}

119. & 洞くつ & 洞窟 & Doukutsu & Cave; ghotto; cavern \\ \cline{1-5}

120. & ばい煙 & 煤煙 & Baien & Soot; smoke \\ \cline{1-5}

121. & 失そう & 失踪 & Shissou & Disappearance \\ \cline{1-5}

122. & 改ざん & 改竄 & Kaizan & Falsification; manipulation; violence \\ \cline{1-5}

123. & えん曲 & 婉曲 & Enkyoku & Euphemism; roundabout \\ \cline{1-5}

124. & けん制 & 牽制 & Kensei & Tow \\ \cline{1-5}

125. & なつ印 & 捺印 & Natsuin & Signet; stigmitization \\ \cline{1-5}

126. & 土のう & 土嚢 & Donou & Sandbag \\ \cline{1-5}

127. & 装てん & 装填 & Souten & Charge; load \\ \cline{1-5}

128. & 抜てき & 抜擢 & Batteki & Selection; choose \\ \cline{1-5}

129. & はく離 & 剥離 & Hakuri & Detachment; coming\slash peeling off; separation \\ \cline{1-5}

130. & はく落 & 剥落 & Hakuraku & To peel off \\ \cline{1-5}

131. & ばく進 & 驀進 & Bakushin & Dash; hurtle \\ \cline{1-5}

132. & ねん挫 & 捻挫 & Nenza & Sprain; twist \\ \cline{1-5}

133. & しゅう恥 & 羞恥 & Shuuchi & Bashfulness; shyness \\ \cline{1-5}

134. & 遮へい & 遮蔽 & Shahei & Shield from \\ \cline{1-5}

135. & 付せん & 付箋 & Fusen & Post-it; tag; docket \\ \cline{1-5}

136. & 容ぼう & 容貌 & Youbou & Looks; appearance; feature; countenance \\ \cline{1-5}

137. & わい曲 & 歪曲 & Waikyoku & Distortion; twists and turns; twist \\ \cline{1-5}

138. & 落いん & 落胤 & Rakuin & The son\slash daughter of (literary); nobleman's illegitimate child \\ \cline{1-5}

139. & 防じん & 防塵 & Boujin & Dust protector \\ \cline{1-5}

140. & 排せつ & 排泄 & Haisetsu & Excrement \\ \cline{1-5}

141. & 殺りく & 殺戮 & Satsuriku & Slaughter; genocide; carnage \\ \cline{1-5}

142. & こん倒 & 昏倒 & Kontou & To faint unconscious \\ \cline{1-5}

143. & ばん回 & 挽回 & Bankai & Recovery; retrieve; restoration \\ \cline{1-5}

144. & 親せき & 親戚 & Shinseki & Relatives \\ \cline{1-5}

145. & 平たんな & 平坦な & Heitan na & Flat; level; even \\ \cline{1-5}

146. & 風ぼう & 風貌 & Fuubou & Feature; countenance \\ \cline{1-5}

147. & ひな形 & 雛形 & Hinagata & Model; form \\ \cline{1-5}

148. & ごう慢な & 傲慢な & Gouman na & Arrogant \\ \cline{1-5}

149. & こう着 & 膠着 & Kouchaku & Adhesion; agglutination; stalemate \\ \cline{1-5}

150. & かん腸 & 浣腸 & Kanchou & Enema \\ \cline{1-5}

151. & 炭そ菌 & 炭疽菌 & Tansokin & Anthrax \\ \cline{1-5}

152. & りょう線 & 稜線 & Ryousen & Mountain ridge \\ \cline{1-5}

153. & 悲そうな & 悲愴な & Hisou na & Sorrowful; mournful \\ \cline{1-5}

154. & えん世 & 厭世 & Ensei & Pessimism; misanthropy \\ \cline{1-5}

155. & がい骨 & 骸骨 & Gaikotsu & Skeleton \\ \cline{1-5}

156. & かい書 & 楷書 & Kaisho & Standard\slash block style \\ \cline{1-5}

157. & 刺しゅう & 刺繍 & Shishuu & Embroidery \\ \cline{1-5}

158. & けん怠 & 倦怠 & Kentai & Ennui; boredom \\ \cline{1-5}

159. & 卑わいな & 卑猥な & Hiwai na & Obscene; vulgar; profane \\ \cline{1-5}

160. & 標ぼう & 標榜 & Hyoubou & Profress \\ \cline{1-5}

161. & へん平 & 扁平 & Henpei & Flat; flat-footed; depressed \\ \cline{1-5}

162. & へき地 & 僻地 & Hekichi & Remote place; out-of-the-way place; isolated district; backwater \\ \cline{1-5}

163. & 流ちょうな & 流暢な & Ryuuchou na & Fluent; facile; smooth \\ \cline{1-5}

164. & きゅう舎 & 厩舎 & Kyuusha & Riding stable \\ \cline{1-5}

165. & 編さん & 編纂 & Hensan & Proofreading; editing; compilation; cutting \\ \cline{1-5}

166. & 風さい & 風采 & Fuusai & Appearance; look; figure \\ \cline{1-5}

167. & こん身 & 渾身 & Konshin & With all one's might \\ \cline{1-5}

168. & 飛しょう体 & 飛翔体 & Hishoutai & Flying object \\ \cline{1-5}

169. & 残がい & 残骸 & Zangai & Wreckage; ruin; wreck \\ \cline{1-5}

170. & 敏しょうな & 敏捷な & Binshou na & Lithe; prompt; nimble; speedy \\ \cline{1-5}

171. & 集じん & 集塵 & Shuujin & Collecting dust \\ \cline{1-5}

172. & 変ぼう & 変貌 & Henbou & Transfiguration; transformation \\ \cline{1-5}

173. & さい銭 & 賽銭 & Saisen & Money offered at a Shinto shrine \\ \cline{1-5}

174. & かく乱 & 撹乱 & Kakuran & Disturbance; perturb \\ \cline{1-5}

175. & 謙そん & 謙遜 & Kenson & Humility; humbleness; modisty \\ \cline{1-5}

176. & 覚せい剤 & 覚醒剤 & Kakuseizai & Stimulant \\ \cline{1-5}

\end{ltabulary}
     