    
\chapter{The Particle どころ}

\begin{center}
\begin{Large}
第269課: The Particle どころ 
\end{Large}
\end{center}
 
\par{ This lesson is all about the easy particle どころ, but don't confuse it with the noun ところ. }
      
\section{The Adverbial Particle どころ}
 
\par{ どころではない means "not possible to do\dothyp{}\dothyp{}\dothyp{}" or "out of the question" and is seen after the 連体形 or a noun. 騒ぎ and 話 are often added after どころ. }

\par{1. 電気が止まれば、寒いどころの話ではないよ。 \hfill\break
If the electricity will stop, then it won't be possible to be cold. }

\par{2. 休暇どころの騒ぎではないよ。 \hfill\break
There's no way I could take the trouble of a vacation. }

\par{3. 忙しくて食事どころではなかったんだ。 \hfill\break
I was too busy to even take a meal. }

\par{4. 忙しくて買い物に行くどころじゃなかった。 \hfill\break
I was so busy that I couldn't go shopping. }

\par{5. 事故で休暇どころの話じゃないよ。 \hfill\break
Due to the accident, talk of vacation is out of the question. }

\par{6. 受験を前にして花見どころの騒ぎではないです。 \hfill\break
Going cherry blossom viewing is out of the question before taking the exam. }

\par{7. まじめどころの騒ぎじゃないさ。 \hfill\break
Serious is out of the question. }

\par{8. 寒いどころじゃないぜ。凍え死ぬぞ！ \hfill\break
Cold is out of the question! I'll freeze to death! }

\par{9. 政治家は正直どころの騒ぎではありません。 \hfill\break
Politicians being honest is out of the question. }

\par{10. 試験を受けないといけないから旅行どころの騒ぎではない。 \hfill\break
Because I have to take an exam, traveling is out of the question. }

\par{11. 天気が悪くては海で泳ぐどころの騒ぎじゃない。 \hfill\break
The weather is bad, so swimming in the ocean is out of the question. }

\par{12. あいつは紳士どころじゃないよ！ \hfill\break
He's far from a gentleman! }

\par{\textbf{The Conjunctive Particle どころか } }

\par{The conjunctive combination particle どころか means "let alone", "far from", etc. It may also be seen as どころではなく. }

\par{13. それどころか、ひどいやつだ。 \hfill\break
Rather, he's a terrible guy. }

\par{14. 教科書には満足どころか全く不満だ。 \hfill\break
The textbook is far from being satisfactory. }

\par{15. いや、それどころか喜んでいます。 \hfill\break
No, rather, I'm glad. }

\par{16. 弟は全然勉強しない。それどころか、１日中ゲームばかりしてる。 \hfill\break
My little brother never studies. Instead, he plays games all day. }

\par{17. 一万円どころか一円も持っていない。 \hfill\break
I don't have a single yen, much less 10,000. }

\par{18. 雨はやむどころではなくいよいよ酷くなった。 \hfill\break
The rain got worse rather than stopping. }

\par{19. それどころか、かなり裕福だそうです。 \hfill\break
On the contrary, that say that they are quite affluent. }

\par{20. 口腔手術を受けて、食べるどころか何も飲めないんだ。 \hfill\break
I had oral surgery, so I can't drink anything and I'm far from eating. }
    