    
\chapter{せめて}

\begin{center}
\begin{Large}
第178課: せめて 
\end{Large}
\end{center}
 
\par{ This really isn't that hard. So, take this as a break from hard studying. }
      
\section{せめて}
 
\par{ せめて means "at the very least". せめて is used when implied that more is desirable. After all, "it's the very least". Because of this, it is often used with the particle くらい・ぐらい to make the statement less specific\slash direct. With は, it may be translated as "if it is unavoidable". }

\par{1. せめて夢の中なら、何でも自由にできるでしょう。 \hfill\break
Shouldn't you at least be able to do anything freely inside a dream? }

\par{2. せめて酒はやめなよ。(Casual) \hfill\break
Quit drinking at the very least. }

\par{3. 小さな事業でも興したいので、せめてパソコンくらいはほしいです。 \hfill\break
Because I want to start up a small business, I want at least a computer. }

\par{4. どんなに経済的にも大変でも、せめて光熱費を払ったほうがいい。 \hfill\break
No matter how financially troubled you are, it's best to at least pay your utility bills. }

\par{5. せめて一万円は支払わなきゃいけねー。（Slang) \hfill\break
If it cannot be helped, I'll have to pay 10,000 yen. }

\par{6. せめて宿題ぐらいはしといてくれ。(Casual; potentially rude; 男性語) \hfill\break
At the very most, do your homework beforehand! }

\par{7. せめて一年間ぐらい日本に住んでみたいんですがねえ。(20年代後半以降) \hfill\break
I'd like to at least live in Japan for a year. }

\par{8. どんなに疲れていても、せめて三十分くらいは勉強してください。 \hfill\break
No matter how tired you are, at least study for thirty minutes. }

\par{9. せめて週に一回くらいは運動した方がいいですよ。 \hfill\break
It's best to at least exercise once a week. }

\par{10. どんなにあの人が嫌いでも、せめて挨拶ぐらいはしたほうがいい。 \hfill\break
No matter how much you hate that person, it's best to at least great him. }

\par{11. 毎日せめて一時間ぐらい勉強してきてほしいですねえ。 \hfill\break
I wish that they'd at least come to study an hour daily. }

\par{12. アメリカでは、せめて一度はハワイへ行ってみたいと思う人が多い。 \hfill\break
There are a lot of people in America who would like to visit Hawaii at least once. }

\par{\textbf{漢字 Note }: ハワイ may also rarely be spelled in 漢字 with the あて字 spelling 布哇. If you are to ever type this, do not use 羽合, which has the same pronunciation but is a town in Japan. }
    